---
title: "Get "
author: "You!"
date: "`r format(Sys.time(), '%m/%d/%Y')`"
format:
    html:
        toc: true
        toc_float: true
        toc_depth: 5
        number_sections: false
        code-fold: true
        code-link: true
        df-print: kable
        embed-resources: true
        self_contained: true
execute:
    cache: true
knitr:
    opts_chunk:
        cache.lazy: false
        cache.vars: ""
        warning: false
        message: false
lightbox: true
---

```{r setup}
#| include: FALSE
library(rrrSingleCellUtils)
library(Seurat)
library(tidyverse)
theme_set(theme_bw())
theme_update(plot.title = element_text(hjust = 0.5))
library(gridExtra)
library(patchwork)

set.seed(1337)
```

# Load in functions

{{< include child_qmds/helper_functions.qmd >}}

# Set up directory structure
```{bash mkdirs}
for directoryName in \
    output \
    output/counts \
    output/degs \
    output/gsea \
    output/figures \
    output/figures/volcano_plots \
    output/figures/gsea \
    output/seurat_objects/individual \
    output/seurat_objects/merged \
    input \
    input/fastqs \
    custom_ref
do
    if [ ! -d ${directoryName} ]
    then
        mkdir -p ${directoryName}
    fi
done
```

## Check required software is available
```{r h5_check}
#| cache: false
hdf5_check()

seurat_version_check()
```

Control: do this on 10x flex single cell data

potential datasets for validation

- GSE246498
    - ZsGreen in tumor cells
    - I think this has no tumor cells

**- GSE247228 -- this one will work**
    - lentiviral green fluorescent protein (GFP) in tumor cells
    - data looks like it will work
    - https://pmc.ncbi.nlm.nih.gov/articles/PMC10830842/
    - realigned to reference, identified tumor cells, and saved as qs object

**- GSE201615**
    - Cas9 in tumor cells, also dsRED
    - To generate sarcoma cells, mesenchymal cells were transduced for the stable expression of Ccne1 and red fluorescent protein
    - https://pmc.ncbi.nlm.nih.gov/articles/PMC9249098/
    - data looks good, wonder if low UMI counts
    - Realigned and made objects, think I know which are tumor cells and saved as qs object

- GSE158700
    - Cas9 and EGFP
    - https://pmc.ncbi.nlm.nih.gov/articles/PMC8168823/
    - this may be a bit tough as the tumors were induced by KO

*- GSE215239*
    - GFP and luciferase in tumor cells
    - no paper yet
    - has no tumor controls
    - likely would work, needs testing
    - Currently making reference for this with GFP and luciferase

- GSE189856
    - pEGFP in some tumor cells, I think?
    - https://pmc.ncbi.nlm.nih.gov/articles/PMC10468402/#Sec11
    - not sure if it's good

- *GSE215394*
    - GFP in tumor cells
    - https://pmc.ncbi.nlm.nih.gov/articles/PMC10951347/
    - meh? All cells seem bunched, only tumor and macs?
    - downloading fastqs right now

- GSE232335 <- gfp normal cells
    - I don't think there are tumor cells in the data

maybe human data to use:
tumor[title] AND "homo sapiens"[orgn] AND ("single cell" OR "singlecell" OR "single-cell") AND "10x" NOT "sorted"
- GSE253013

Make fasta
make gtf

Make new reference with cat of mouse and tumor markers

Align to new ref


# Read in raw data and process to Seurat objects
The metadata for all samples is stored in misc/sample_metadata.tsv.

Note that in the sample_metadata, for samples with multiple references, the cutoff format is "GRCh38-40000;mm10---20000" where the species_pattern is at the beginning of the string, followed by the cutoff value (no dash or anything). This is split from the second cutoff value by a semicolon. If there is only one reference, just use the cutoff value, no species pattern is needed.

{{< include child_qmds/read_raw_data.qmd >}}

# Merge the data by species and make DimPlots

{{< include child_qmds/merge_dimplots.qmd >}}

# Run differential expression and GSEA

{{< include child_qmds/degs_gseaplots.qmd >}}


# Session info
```{r session_info}
sessionInfo()
```
