---
title: "Validate that scanBit Correctly Identifies Tumor Cells"
author: "Matt Cannon and Matt Gust"
date: "`r format(Sys.time(), '%m/%d/%Y')`"
format:
    html:
        toc: true
        toc_float: true
        toc_depth: 5
        number_sections: false
        code-fold: true
        code-link: true
        df-print: kable
        embed-resources: true
        self_contained: true
execute:
    cache: true
knitr:
    opts_chunk:
        cache.lazy: false
        cache.vars: ""
        warning: false
        message: false
lightbox: true
---

```{r setup}
#| include: FALSE
library(rrrSingleCellUtils)
library(Seurat)
library(tidyverse)
theme_set(theme_bw())
theme_update(
    plot.title = element_text(hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5)
)
library(gridExtra)
library(patchwork)
library(scanBit)
library(qs2)

set.seed(1337)
```

# Load in helper R functions
{{< include child_qmds/helper_functions.qmd >}}

# Set up directory structure
```{bash mkdirs}
for directoryName in \
    /gfpf0/scratch/mvc002/scanbit \
    input/10x_flex \
    output \
    slurmOut \
    output/counts \
    output/degs \
    output/gsea \
    output/refs \
    output/figures \
    output/figures/qc_cutoffs \
    output/figures/volcano_plots \
    output/figures/gsea \
    output/seurat_objects/individual \
    output/seurat_objects/merged \
    output/read_depth/depths/bcf_depth \
    output/read_depth/depths/call_count \
    output/downsample/scanbit_out/x_samples/ \
    output/read_depth/scanbit_out/x_samples \
    output/read_depth/depths \
    output/snv/tumor/x3 \
    output/snv/human/ \
    output/snv/flex \
    output/snv/mouse \
    output/snv/transgene/ \
    output/snv_processing/norm_mouse/ \
    input \
    input/fastqs \
    custom_ref
do
    if [ ! -d ${directoryName} ]
    then
        mkdir -p ${directoryName}
    fi
done
```

## Check required software is available
```{r h5_check}
#| cache: false
hdf5_check()

seurat_version_check()
```

# For these analysis we are going to do some key validations and analyses.

- Run variant calling on 10x flex data to quantify the sequencing error rate at varying read depths.
    - Data from https://www.10xgenomics.com/datasets/40k-mixture-of-cells-dissociated-from-4-fixed-mouse-tissues-using-manual-dissociation-multiplexed-samples-4-probe-barcodes-1-standard
- Analyze samples where the tumor cells express a marker that allows us to know which cells should be called as tumor and then assess accuracy of the method.
    - S0105 - K7M2-RFP tibia tumor
    - S0122 - K7M2 sLPmCherry lung mets 10 days post injection. Flow sorted Cherry positive fraction 1:1:1 mix of CD45+, Epcam+ and double negative
    - GSE247228 - lentiviral green fluorescent protein (GFP) in melanoma cells
    - GSE201615 - Cas9 in soft-tissue sarcoma cells, also dsRED
    - S0260 - barcoded F420 tumor samples
    - S0261 - barcoded F420 tumor sample
- Analyze samples from genetically distinct mouse strains where we know how the clustering should look.
    - S0066 - B6 lung
    - S0074 - Balb/c lung
- Analyze mouse OS models where the tumor cells derive from the host mouse strain to simulate a genetic divergence more expected in patients.
    - S0066 - (control) B6 lung
    - S0067 - B6, F420 injected, lungs from ROB 273 A, B, D; Epcam postive:negative 1:1
    - S0068 - B6, F420 injected, lungs from ROB 273E and ROB 274A; Epcam postive:negative 1:1
    - S0074 - (control) Balb/c lung
    - S0075 - Balb/c, K7M2 injected, lungs; Epcam postive:negative 1:1
    - S0105 - K7M2-RFP tibia tumor
    - S0200 - (control) F420 cellculture p17
    - S0201 - (control) K7M2 cellculture p17
    - S0246 - F420 lung mets and niche cells ctrl
    - S0259 - (control) F420, barcoded, preinjection t0 ctrl
    - S0260 - barcoded F420 tumor samples
    - S0261 - barcoded F420 tumor sample
 - Assess scanBit results when run on human patient tumor samples.
    - GSE152048 OS samples
    - GSE162454 OS samples
    - GSE210358 prostate cancer samples
    - GSE222703 renal cell carcinoma samples
    - GSE169396 bone samples
    - GSE237579 healthy PBMC samples

# Download raw data from GEO/SRA, process it and re-align Roberts lab data to GFP/RFP reference
This also downloads the flex data from 10x

The output here will be cellrange counts folders in output/cellranger_out/

{{< include child_qmds/process_raw_data.qmd >}}

# Read in raw data and process to Seurat objects
The metadata for all samples is stored in misc/sample_metadata.tsv.

Note that in the sample_metadata, for samples with multiple references, the cutoff format is "GRCh38-40000;mm10---20000" where the species_pattern is at the beginning of the string, followed by the cutoff value (no dash or anything). This is split from the second cutoff value by a semicolon. If there is only one reference, just use the cutoff value, no species pattern is needed.

{{< include child_qmds/make_Seurat_objs.qmd >}}

# Run scanBit on the Seurat objects

{{< include child_qmds/call_snvs.qmd >}}

# Merge the data by species and make DimPlots

##{{< #include child_qmds/merge_dimplots.qmd >}}

# Test scanBit on prostate cancer
This is a tumor type with low mutational burden

##{{< #include child_qmds/prostate_cancer.qmd >}}



# Session info
```{r session_info}
sessionInfo()
```
