---
title: "Validate that scanBit Correctly Identifies Tumor Cells"
author: "Matt Cannon and Matt Gust"
date: "`r format(Sys.time(), '%m/%d/%Y')`"
format:
    html:
        toc: true
        toc_float: true
        toc_depth: 5
        number_sections: false
        code-fold: true
        code-link: true
        df-print: kable
        embed-resources: true
        self_contained: true
execute:
    cache: true
knitr:
    opts_chunk:
        cache.lazy: false
        warning: false
        message: false
lightbox: true
---

```{r setup}
#| include: FALSE
#| cache: FALSE
library(rrrSingleCellUtils)
library(Seurat)
library(tidyverse)
theme_set(theme_bw())
theme_update(
    plot.title = element_text(hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5)
)
library(gridExtra)
library(patchwork)
library(scanBit)
library(qs2)
library(HDF5Array)

set.seed(1337)

# Load in general helper R functions
source("analysis_modules/general_helper_functions.R")
```

# Set up directory structure
```{bash mkdirs}
for directoryName in \
    /gpfs0/scratch/mvc002/scanbit \
    input/10x_flex \
    output \
    slurmOut \
    output/01_process_raw/counts \
    output/01_process_raw/figures/qc_cutoffs \
    output/01_process_raw/rdata/individual \
    output/01_process_raw/rdata/merged \
    output/02_snv_calling/rdata/merged \
    output/02_snv_calling/figures \
    output/02_snv_calling/snv/flex \
    output/02_snv_calling/snv/transgene \
    output/02_snv_calling/snv/mouse \
    output/02_snv_calling/snv/human \
    output/03_test_flex/counts/flex_var_counts_d1 \
    output/03_test_flex/counts/mouse \
    output/03_test_flex/figures \
    output/03_test_flex/snvs \
    output/04_read_depth_analysis/snv/geo \
    output/05_mouse_strain/rdata/merged \
    output/10_cmp_other_methods/scevan_results \
    output/10_cmp_other_methods/scatomic_results \
    output/10_cmp_other_methods/copykat_results/temp \
    output/10_cmp_other_methods/copykat_results/rdata \
    input \
    input/fastqs \
    custom_ref
do
    if [ ! -d ${directoryName} ]
    then
        mkdir -p ${directoryName}
    fi
done
```

## Check required software is available
```{r h5_check}
#| cache: false
#hdf5_check()

seurat_version_check()
```

# For these analysis we are going to do some key validations and analyses.

- Run variant calling on 10x flex data to quantify the sequencing error rate at varying read depths.
    - Data from https://www.10xgenomics.com/datasets/40k-mixture-of-cells-dissociated-from-4-fixed-mouse-tissues-using-manual-dissociation-multiplexed-samples-4-probe-barcodes-1-standard
- Analyze samples where the tumor cells express a marker that allows us to know which cells should be called as tumor and then assess accuracy of the method.
    - S0105 - K7M2-RFP tibia tumor
    - S0122 - K7M2 sLPmCherry lung mets 10 days post injection. Flow sorted Cherry positive fraction 1:1:1 mix of CD45+, Epcam+ and double negative
    - GSE247228 - lentiviral green fluorescent protein (GFP) in melanoma cells
    - GSE201615 - Cas9 in soft-tissue sarcoma cells, also dsRED
    - S0260 - barcoded F420 tumor samples
    - S0261 - barcoded F420 tumor sample
- Analyze samples from genetically distinct mouse strains where we know how the clustering should look.
    - S0066 - B6 lung
    - S0074 - Balb/c lung
- Analyze mouse OS models where the tumor cells derive from the host mouse strain to simulate a genetic divergence more expected in patients.
    - S0066 - (control) B6 lung
    - S0067 - B6, F420 injected, lungs from ROB 273 A, B, D; Epcam postive:negative 1:1
    - S0068 - B6, F420 injected, lungs from ROB 273E and ROB 274A; Epcam postive:negative 1:1
    - S0074 - (control) Balb/c lung
    - S0075 - Balb/c, K7M2 injected, lungs; Epcam postive:negative 1:1
    - S0105 - K7M2-RFP tibia tumor
    - S0200 - (control) F420 cellculture p17
    - S0201 - (control) K7M2 cellculture p17
    - S0246 - F420 lung mets and niche cells ctrl
    - S0259 - (control) F420, barcoded, preinjection t0 ctrl
    - S0260 - barcoded F420 tumor samples
    - S0261 - barcoded F420 tumor sample
 - Assess scanBit results when run on human patient tumor samples.
    - GSE152048 OS samples
    - GSE162454 OS samples
    - GSE210358 prostate cancer samples
    - GSE222703 renal cell carcinoma samples
    - GSE169396 bone samples
    - GSE237579 healthy PBMC samples

# Download raw data from GEO/SRA, process it and re-align Roberts lab data to GFP/RFP reference
This also downloads the flex data from 10x

The output here will be cellrange counts folders in output/cellranger_out/

The metadata for all samples is stored in misc/sample_metadata.tsv.

Note that in the sample_metadata, for samples with multiple references, the cutoff format is "GRCh38-40000;mm10---20000" where the species_pattern is at the beginning of the string, followed by the cutoff value (no dash or anything). This is split from the second cutoff value by a semicolon. If there is only one reference, just use the cutoff value, no species pattern is needed.

{{< include analysis_modules/01_process_raw/01_1_process_raw_data.qmd >}}

# Run scanBit on the Seurat objects

{{< include analysis_modules/02_snv_calling/02_1_call_snvs.qmd >}}

# Run analysis on flex data to calculate accuracy

{{< include analysis_modules/03_test_flex/03_1_test_flex.qmd >}}

# Downsample GEO samples to one cell and quantify read depth coverage

{{< include analysis_modules/04_read_depth/04_1_read_depth_analysis.qmd >}}

# Look at how well scanBit works to distinguish B6 from Balbc

include child_qmds/test_mouse_strains.qmd >}}

# Look at how well scanBit can distinguish B6 from F420 and Balbc from K7M2
This is more similar to the situation found in patients where the tumor cells derive from the patient

include child_qmds/F420_K7M2_analysis.qmd >}}

# Downsample each sample to a range of cells per cluster
This is to test the minimum number of cells/reads needed to accurately identify genetically distinct populations.

Also, I will use the data where I downsample to one cell to calculate coverage provided by one cell on average

include child_qmds/downsample_analysis.qmd >}}

# Look at samples from tumors with transgene markers

include child_qmds/evaluate_tumor_markers.qmd >}}

# Look at human patient data

include child_qmds/human_validation.qmd >}}

# Quantify average run time and RAM usage

include child_qmds/run_time_analysis.qmd >}}

# Run SCEVAN and other alternative methods
{{< include analysis_modules/10_cmp_other_methods/10_1_scevan.qmd >}}

# Session info
```{r session_info}
sessionInfo()
```
