## Run scanBit on individual cells with min_read_depth of 1 to use for read depth and accuracy analysis
For this I'm keeping ten individual cells per cluster, and each cell is assigned to its own cell_group so that I am calculating read depth data for many individual cells.
```{r readdepth_scanbit}
samples_geo <-
    read_tsv("misc/validation_geo_metadata.tsv", show_col_types = FALSE) %>%
    filter(grepl("tumor validation|healthy samples", sample_group)) %>%
    pull(Sample_ID) %>%
    unique()

n_cell_cutoffs <- 1

# Keeping mc.cores low since each run submits 100+ jobs
output <-
    parallel::mclapply(
        samples_geo,
        mc.cores = 2,
        mc.preschedule = FALSE,
        function(this_sample) {
            sobj <-
                qs::qread(paste0(
                    "output/seurat_objects/individual/",
                    this_sample,
                    "_human.qs"
                ))

            sobj$cell_group <-
                paste0(
                    "clust_",
                    sobj$individ_clusters,
                    "_",
                    sobj$cell_barcode
                )

            Idents(sobj) <- sobj$individ_clusters
            set.seed(1337)
            sobj <- subset(sobj, downsample = 10)

            Idents(sobj) <-
                sobj$cell_group

            scanbit_downsample_loop(
                sobj,
                cutoff_vector = n_cell_cutoffs,
                min_read_depth = c(1, 2, 5, 10, 20),
                out_dir_base = file.path(
                    "output/read_depth/scanbit_out/geo_samples/",
                    this_sample
                ),
                keep_temp_dir = TRUE,
                sample_label = this_sample,
                ploidy = "GRCh38",
                ref_fasta = "/home/gdrobertslab/lab/GenRef/10x-hg38/fasta/genome.fa"
            )
        }
    )
```

## Batch job to count site coverage of GEO human patient samples
This includes any site with coverage, not just variant calls
```{bash readdepth_count_depth}
sbatch sbatchCmds/geo_bcf_depths.sh
```

## Calculate stats on nCount_RNA for single cell data
```{r readdepth_ncount_rna_stats}
#| dependson: [readdepth_scanbit]
list.files(
    path = "output/read_depth/scanbit_out/geo_samples",
    pattern = "cell_barcode_table_.+_1_cells.txt",
    recursive = TRUE,
    full.names = TRUE
) %>%
    read_tsv(
        id = "sample",
        show_col_types = FALSE
    ) %>%
    summarize(
        mean_ncount = mean(nCount_RNA),
        median_ncount = median(nCount_RNA),
        min_ncount = min(nCount_RNA),
        max_ncount = max(nCount_RNA)
    ) %>%
    write_tsv("output/read_depth/ncount_summary.tsv")
```

## Calculate the # of positions covered where any cell has a variant
```{r readdepth_var_positions}
#| cache.vars: [var_depths, ave_var_depths]
all_var_files <-
    list.files(
        path = "output/read_depth/depths/bcf_depth",
        pattern = "variant_pos",
        full.names = TRUE
    )

var_depths <-
    read_tsv(
        all_var_files,
        id = "cell_sample",
        show_col_types = FALSE,
        skip_empty_rows = FALSE
    ) %>%
    mutate(
        cell_sample = basename(cell_sample) %>%
            str_remove("variant_pos_") %>%
            str_remove(".txt"),
        group = "Variant sites"
    )

ave_var_depths <-
    var_depths %>%
    pivot_wider(
        names_from = "depth",
        values_from = "count",
        values_fill = 0
    ) %>%
    pivot_longer(
        cols = c(-cell_sample, -group),
        names_to = "depth",
        values_to = "count"
    ) %>%
    group_by(depth) %>%
    summarize(
        mean_count_var = mean(count),
        median_count_var = median(count),
        se_count = sd(count) / sqrt(n()),
        .groups = "drop",
    ) %>%
    mutate(depth = as.numeric(depth)) %>%
    arrange(desc(depth)) %>%
    mutate(cum_count_var = cumsum(mean_count_var)) %>%
    arrange(depth)

write_tsv(
    ave_var_depths,
    "output/read_depth/n_var_count_summary.tsv"
)
```

## Compile all single-cell variant call depths
```{r readdepth_compile_depths}
#| cache.vars: [all_depths, ave_depths]
all_depth_files <-
    list.files(
        path = "output/read_depth/depths/bcf_depth",
        pattern = "bcf_depth_G",
        full.names = TRUE
    )
all_depths <-
    read_tsv(
        all_depth_files,
        id = "cell_sample",
        show_col_types = FALSE,
    ) %>%
    mutate(
        cell_sample = basename(cell_sample) %>%
            str_remove("bcf_depth_") %>%
            str_remove(".txt"),
        group = "All sites"
    )

ave_depths <-
    all_depths %>%
    filter(depth != 1) %>%
    pivot_wider(
        names_from = "depth",
        values_from = "count",
        values_fill = 0
    ) %>%
    pivot_longer(
        cols = c(-cell_sample, -group),
        names_to = "depth",
        values_to = "count"
    ) %>%
    group_by(depth) %>%
    summarize(
        mean_count = mean(count),
        median_count_var = median(count),
        se_count = sd(count) / sqrt(n()),
        .groups = "drop"
    ) %>%
    mutate(depth = as.numeric(depth)) %>%
    arrange(desc(depth)) %>%
    mutate(cum_count_var = cumsum(mean_count)) %>%
    arrange(depth)

write_tsv(
    ave_depths,
    "output/read_depth/n_covered_count_summary.tsv"
)
```

```{r readdepth_plot_coverage}
site_covered_plot <-
    ggplot() +
    geom_point(
        data = ave_depths %>% filter(depth <= 30),
        aes(x = depth, y = mean_count),
        color = "black"
    ) +
    geom_point(
        data = ave_var_depths %>% filter(depth <= 30),
        aes(x = depth, y = mean_count_var),
        color = "red"
    ) +
    scale_y_log10(labels = scales::label_comma()) +
    labs(
        title = "Mean Sites Covered to a Given Depth",
        subtitle = "Red: mean variable sites; Black: mean total sites",
        x = "Depth",
        y = "Average number of positions covered"
    )

ggsave(
    "output/figures/mean_coverage_1_cell.pdf",
    site_covered_plot,
    width = 6,
    height = 4
)

set.seed(1337) # For jitter
jitter_coverage_plot <-
    rbind(all_depths, var_depths) %>%
    group_by(group, cell_sample) %>%
    arrange(desc(depth), .by_group = TRUE) %>%
    mutate(cum_count = cumsum(count)) %>%
    arrange(depth, .by_group = TRUE) %>%
    ungroup() %>%
    filter(depth <= 20 & depth != 1) %>%
    select(-count) %>%
    pivot_wider(
        names_from = "depth",
        values_from = "cum_count",
        values_fill = 0
    ) %>%
    pivot_longer(
        cols = c(-cell_sample, -group),
        names_to = "depth",
        values_to = "count"
    ) %>%
    mutate(
        depth_factor = as.factor(depth) %>%
            fct_reorder(as.numeric(depth))
    ) %>%
    ggplot(
        aes(
            x = depth_factor,
            y = count,
            color = group
        )
    ) +
    geom_jitter(height = 0, alpha = 0.5, size = 0.5) +
    scale_y_log10(labels = scales::label_comma()) +
    scale_color_manual(values = plot_cols[1:2]) +
    labs(
        title = "Number of Total and Non-reference Sites Covered by Single Cells Using Varying Cutoffs",
        x = "Read depth cutoff",
        y = "Number of sites covered when using read-depth cutoff of X",
        color = NULL
    ) +
    theme(
        legend.position = c(0.9, 0.9)
    )

ggsave_wrap(
    file_stub = "output/figures/total_and_nonref_site_coverage",
    file_types = c("png", "pdf"),
    jitter_coverage_plot,
    width = 8,
    height = 6
)

#rm(all_depths, var_depths, ave_var_depths)
```

## Make a plot of the percentage of calls that are variant vs depth
```{r readdepth_perc_variant_plot}
all_depth_data <-
    full_join(
        all_depths %>%
            rename(total_depth = count) %>%
            select(-group) %>%
            filter(depth != 1),
        var_depths %>%
            rename(variant_depth = count) %>%
            select(-group)
    ) %>%
    mutate(
        variant_depth = replace_na(variant_depth, 0),
        perc_variant = variant_depth / total_depth,
        depth_fct = as.factor(depth) %>% fct_reorder(depth)
    ) %>%
    filter(depth <= 30)

all_depth_data %>%
    group_by(depth) %>%
    summarize(
        mean_perc_variant = mean(perc_variant) * 100,
        pct_no_variants = mean(variant_depth == 0) * 100,
        .groups = "drop"
    ) %>%
    arrange(depth) %>%
    write_tsv("output/read_depth/perc_variant_summary.tsv")

perc_variant_plot <-
    all_depth_data %>%
    ggplot(
        aes(
            x = depth_fct,
            y = perc_variant
        )
    ) +
    geom_jitter(alpha = 0.8, size = 0.5) +
    scale_y_log10(labels = scales::label_percent()) +
    labs(
        title = "Percent of Variant Calls by Read Depth",
        x = "Read Depth",
        y = "Percent of Calls That Are Variant"
    )
ggsave(
    "output/figures/percent_variant_calls_1_cell.pdf",
    perc_variant_plot,
    width = 6,
    height = 4
)

perc_no_variants_plot <-
    all_depth_data %>%
    group_by(depth_fct) %>%
    summarize(pct_no_variants = mean(variant_depth == 0), .groups = "drop") %>%
    ggplot(
        aes(
            x = depth_fct,
            y = pct_no_variants
        )
    ) +
    geom_col() +
    scale_y_continuous(labels = scales::label_percent()) +
    labs(
        title = "Percent of Cells With No Variant Calls by Read Depth",
        x = "Read Depth",
        y = "Percent of Cells With No Variant Calls"
    )
ggsave(
    "output/figures/percent_no_variant_calls_1_cell.pdf",
    width = 6,
    height = 4
)
```
## Calculate expected number of mistakes for each depth
Based on flex accuracy calculations
```{r readdepth_accuracy_calc}
#| cache.vars: [expected_mis_calls]
# Make a named vector or accuracy - name is depth, value is % accurate
flex_accuracy_prop <-
    read_tsv("output/flex/flex_accuracy.txt", show_col_types = FALSE) %>%
    pull(mean_perc_wrong, name = depth) / 100

expected_mis_calls <-
    ave_depths %>%
    mutate(exp_bad_variants = mean_count * flex_accuracy_prop[depth])

expected_errors_plot <-
    expected_mis_calls %>%
    head(n = 100) %>%
    ggplot() +
    geom_point(aes(x = depth, y = mean_count), color = "black") +
    geom_point(
        aes(
            x = depth,
            y = exp_bad_variants
        ),
        color = "red",
        size = 2
    ) +
    scale_y_log10(labels = scales::label_comma())

ggsave(
    "output/figures/expected_errors_1_cell.png",
    expected_errors_plot,
    width = 6,
    height = 4
)
```

## Count the number of variant (non-reference) calls made in single cells
```{bash readdepth_actual_n_vars}
sbatch sbatchCmds/count_single_cell_variant_calls.sh
```

## Calculate the average number of variants actually called by depth
```{r readdepth_compile_depths_ave}
all_call_files <-
    list.files(
        path = "output/read_depth/depths/call_count",
        pattern = "bcf_calls_depth_G0",
        full.names = TRUE
    )
all_call_depths <-
    read_tsv(
        all_call_files,
        id = "cell_sample",
        show_col_types = FALSE,
        col_names = c("chr", "alt", "depth")
    ) %>%
    mutate(
        cell_sample = basename(cell_sample) %>%
            str_remove("bcf_calls_depth_") %>%
            str_remove(".txt")
    ) %>%
    select(-alt, -chr)

average_num_calls <-
    all_call_depths %>%
    group_by(depth, cell_sample) %>%
    summarize(n_vars = n(), .groups = "drop") %>%
    pivot_wider(
        names_from = "depth",
        values_from = "n_vars",
        values_fill = 0
    ) %>%
    pivot_longer(
        cols = c(-cell_sample),
        names_to = "depth",
        values_to = "count"
    ) %>%
    group_by(depth) %>%
    mutate(depth = as.numeric(depth)) %>%
    summarize(mean_n_vars = mean(count)) %>%
    arrange(depth)

n_vars_plot <-
    average_num_calls %>%
    filter(depth <= 30) %>%
    ggplot() +
    geom_point(aes(x = depth, y = mean_n_vars), color = "black") +
    labs(
        title = "Average Number of Variants Called by Depth",
        subtitle = "In Variant Call Data on Single Cells",
        x = "Depth",
        y = "Mean number of variants called per cell"
    )

ggsave(
    "output/figures/average_vars_1_cell.pdf",
    n_vars_plot,
    width = 6,
    height = 4
)

calls_table <-
    full_join(expected_mis_calls, average_num_calls, by = "depth") %>%
    filter(depth != 1)

calls_summary_plot <-
    calls_table %>%
    rename(wrong_variants = exp_bad_variants, called_variants = mean_n_vars) %>%
    pivot_longer(
        cols = c(wrong_variants, called_variants),
        names_to = "var_type",
        values_to = "count"
    ) %>%
    mutate(
        var_type = str_replace(var_type, "_", " ") %>%
            str_to_title()
    ) %>%
    filter(depth <= 30) %>%
    ggplot(aes(x = depth, y = count, group = depth)) +
    geom_line() +
    geom_point(aes(color = var_type, shape = var_type), size = 2) +
    scale_color_manual(values = c("black", "red"), name = NULL) +
    scale_shape_manual(values = c(19, 4), name = NULL) +
    labs(
        title = "Total and Inaccurate Variant Calls",
        subtitle = "For single cells from osteosarcoma and bone samples",
        x = "Depth",
        y = "Number of variant calls"
    ) +
    theme(
        legend.position = c(0.85, 0.9),
        legend.background = element_blank()
    )

ggsave(
    "output/figures/wrong_vs_called_nvars_1_cell.pdf",
    calls_summary_plot,
    width = 6,
    height = 4
)

percent_wrong_table <-
    calls_table %>%
    rename(wrong_variants = exp_bad_variants, called_variants = mean_n_vars) %>%
    mutate(perc_wrong = wrong_variants / called_variants) %>%
    arrange(desc(depth)) %>%
    mutate(
        mean_count = replace_na(mean_count, 0),
        called_variants = replace_na(called_variants, 0),
        wrong_variants = replace_na(wrong_variants, 0),
        cum_sites_cov = cumsum(mean_count),
        cum_var_sites = cumsum(called_variants),
        cum_wrong_vars = cumsum(wrong_variants)
    ) %>%
    arrange(depth)

write_tsv(
    percent_wrong_table,
    file = "output/read_depth/percent_wrong_table.tsv"
)

percent_wrong_plot <-
    percent_wrong_table %>%
    filter(depth <= 30) %>%
    ggplot(
        aes(
            x = depth,
            y = perc_wrong
        )
    ) +
    geom_point() +
    scale_y_continuous(labels = scales::label_percent()) +
    labs(
        title = "Percent of Inaccurate Variant Calls by Depth",
        x = "Depth",
        y = "Percent inaccurate variant calls"
    ) +
    theme_bw()

ggsave(
    "output/figures/percent_wrong_1_cell.pdf",
    percent_wrong_plot,
    width = 6,
    height = 4
)
```

## Calculate how many sites have shared variant information per sample
```{bash readdepth_calc_shared}
#| dependson: [readdepth_scanbit]
sbatch sbatchCmds/count_single_cell_shared_sites.sh
```

```{r readdepth_shared_sites}
#| dependson: [readdepth_calc_shared]
shared_files <-
    list.files(
        path = "output/read_depth/shared",
        pattern = "shared.+\\.txt",
        full.names = TRUE
    )

parse_shared <- function(file_name) {
    dataset <- read_tsv(
        file_name,
        name_repair = "minimal",
        show_col_types = FALSE,
        id = "dataset"
    ) |>
        mutate(
            dataset = str_remove(dataset, ".+shared_") |>
                str_remove(".txt")
        ) |>
        pivot_longer(
            cols = starts_with("clust"),
            names_to = "cell_2",
            values_to = "shared_sites"
        ) |>
        filter(cell_1 != cell_2) |>
        separate(dataset, into = c("sample", "cutoff"), sep = "_") %>%
        mutate(cutoff = factor(cutoff, levels = c("1", "2", "5", "10", "20")))

    return(dataset)
}

shared_sites <-
    lapply(
        shared_files,
        parse_shared
    ) %>%
    bind_rows()

pdf(
    "output/figures/shared_sites/shared_sites_heatmaps.pdf",
    width = 10,
    height = 10
)
for (this_sample in unique(shared_sites$sample)) {
    print(shared_sites %>%
        filter(sample == this_sample & cutoff == 1) %>%
        mutate(shared_sites = as.integer(shared_sites)) %>%
        select(cell_1, cell_2, shared_sites) %>%
        pivot_wider(
            names_from = cell_2,
            values_from = shared_sites,
            values_fill = NA
        ) %>%
        column_to_rownames("cell_1") %>%
        as.matrix() %>%
        ComplexHeatmap::pheatmap(
            display_numbers = TRUE,
            fontsize_number = 10,
            number_format =  "%.0f",
            fontsize = 5,
            use_raster = TRUE,
            raster_quality = 4,
            main = paste0(
                "Number of Sites Shared Between Individual Cells for a Subset of Sample ",
                this_sample
            ),
            heatmap_legend_param = list(title = "Shared\nsites")
        ) +
        labs(fill = NULL)
    )
}
dev.off()

shared_sites_histograms <-
    shared_sites %>%
    filter(cutoff != "1") %>%
    ggplot(
        aes(
            x = shared_sites,
            fill = cutoff
        )
    ) +
    geom_histogram(bins = 200) +
    scale_x_continuous(n.breaks = 20) +
    facet_wrap(
        ~ cutoff,
        ncol = 1,
        scales = "free_y"
    ) +
    labs(
        title = "Distribution of Number of Genomic Positions With Shared Variant\nInformation Between Two Cells",
        subtitle = "For sites where either cell has a non-reference variant call",
        x = "Shared covered genomic positions",
        y = "Count",
        fill = "Minimum read depth\nto call variants"
    ) +
    theme(
        legend.position = "none"
    )

ggsave(
    "output/figures/shared_sites/shared_sites_histograms.pdf",
    plot = shared_sites_histograms,
    width = 8,
    height = 5
)

shared_sites %>%
    group_by(cutoff) %>%
    filter(cutoff != "1") %>%
    summarize(
        mean_shared_sites = mean(shared_sites),
        median_shared_sites = median(shared_sites)
    ) %>%
    write_tsv("output/read_depth/shared_sites_stats.tsv")
```
