## Prep the flex data
We're going to run scanBit on a merged object for this so the data all ends up in one BCF file.
```{r scanbit_prep_flex}
#| dependson: read_raw_human
#| cache.vars: flex_cell_barcode_group_table
flex_samples <-
    read_tsv("misc/flex_metadata.tsv", show_col_types = FALSE) %>%
    pull(Sample_ID)

flex_sobj_list <-
    lapply(
        flex_samples,
        function(this_sample) {
            qs::qread(paste0(
                "output/seurat_objects/individual/",
                this_sample,
                "_mouse.qs"
            ))
        }
    )

merged_flex <-
    merge(flex_sobj_list[[1]], flex_sobj_list[-1]) %>%
    process_seurat()

merged_flex$cell_group <-
    paste0(
        merged_flex$cluster_label,
        "_c",
        merged_flex$individ_clusters
    )

flex_dimplot <-
    (DimPlot(
        merged_flex,
        group.by = "Name",
        label = TRUE,
        label.box = TRUE,
        label.size = 2,
        repel = TRUE,
        ncol = 1
    ) +
        NoLegend()) +
    (DimPlot(
        merged_flex,
        group.by = "cell_group",
        label = TRUE,
        label.box = TRUE,
        label.size = 2,
        repel = TRUE,
        ncol = 1
    ) +
        NoLegend()) +
    (DimPlot(
        merged_flex,
        group.by = "cell_type",
        label = TRUE,
        label.box = TRUE,
        label.size = 2,
        repel = TRUE,
        ncol = 1
    ) +
        NoLegend())

ggsave(
    "output/figures/dimplot_flex_cell_type_group.pdf",
    flex_dimplot,
    width = 10,
    height = 20
)

flex_cell_barcode_group_table <-
    merged_flex@meta.data %>%
    select(cell_group, cell_barcode, bam_file) %>%
    as_tibble()
```

## Run scanBit on the flex data
```{r scanbit_run_flex}
#| dependson: scanbit_prep_flex
# This was erroring out because there aren't enough snvs
# I lowered min_snvs_per_cluster to 5 to try to get it to work
snp_tree <-
    get_snp_tree(
        cellid_bam_table = flex_cell_barcode_group_table,
        ploidy = "mm10",
        output_dir = "output/snv/flex/",
        temp_dir = "output/snv/flex/tempdir",
        output_base_name = "flex_snvs",
        ref_fasta = "/reference/mus_musculus/mm10/ucsc_assmebly/illumina_download/Sequence/WholeGenomeFasta/genome.fa",
        min_depth = c(1, 5, 10, 20, 30),
        job_base = "flex_sbatch",
        log_base = "slurmOut/flex_snv",
        min_snvs_per_cluster = 5,
        n_bootstraps = 10000,
        max_prop_missing_at_site = 0.9,
        tree_image_type = "pdf",
        cleanup = FALSE,
        other_job_header_options = c(
            "--time=8:00:00",
            "--partition=himem,general"
        ),
        other_batch_options = c(
            "ml purge",
            "ml miniforge3",
            'eval "$(conda shell.bash hook)"'
        ),
    )
```

## Prep the mouse samples to run scanBit
We're going to merge all of these data together so the output is one big tree.
We'll break it down further later on.
```{r scanbit_prep_mouse}
#| dependson: read_raw_mouse_references
#| cache.vars: [mouse_cell_barcode_group_table]
mouse_samples <-
    read_tsv(
        "misc/mouse_samples_Roberts_metadata.tsv",
        show_col_types = FALSE
    ) %>%
    pull(Sample_ID) %>%
    c(
        read_tsv(
            "misc/mouse_samples_Roberts_realign_metadata.tsv",
            show_col_types = FALSE
        ) %>%
            pull(Sample_ID)
    )

mouse_sobj_list <-
    lapply(
        mouse_samples,
        function(this_sample) {
            qs::qread(paste0(
                "output/seurat_objects/individual/",
                this_sample,
                "_mouse.qs"
            ))
        }
    )
names(mouse_sobj_list) <- mouse_samples

# With the clustering I did initially, S0067 had a cluster that was roughly
# 50/50 tumor normal and it messes up the whole F420/B6 tree
# This is a good example of how important good clustering is to get this to work
# I renamed the original tree plots with _orig.pdf to maybe show in supplement
# Reclustering so this mixed cluster is split fixes the issue

mouse_sobj_list$S0067 <-
    mouse_sobj_list$S0067 %>%
    process_seurat(resolution = 0.25)

mouse_sobj_list$S0067$individ_clusters <- mouse_sobj_list$S0067$seurat_clusters
mouse_sobj_list$S0067$cell_group <-
    paste0(
        mouse_sobj_list$S0067$cluster_label,
        "_c",
        mouse_sobj_list$S0067$individ_clusters
    )

DimPlot(
    mouse_sobj_list$S0067,
    group.by = c("cell_group", "cell_type"),
    label = TRUE,
    label.box = TRUE,
    repel = TRUE
) +
    NoLegend()

merged_mouse <-
    merge(mouse_sobj_list[[1]], mouse_sobj_list[-1]) %>%
    process_seurat()

merged_mouse$cell_group <-
    paste0(
        merged_mouse$cluster_label,
        "_c",
        merged_mouse$individ_clusters
    )

qs::qsave(merged_mouse, "output/seurat_objects/merged/merged_mouse.qs")

mouse_cell_barcode_group_table <-
    merged_mouse@meta.data %>%
    select(cell_group, cell_barcode, bam_file) %>%
    as_tibble()
```

## Run scanBit on the mouse data from our lab
```{r scanbit_run_mouse}
#| dependson: run_scanbit_prep_mouse
snp_tree <-
    get_snp_tree(
        cellid_bam_table = mouse_cell_barcode_group_table,
        ploidy = "mm10",
        output_dir = "output/snv/mouse/",
        temp_dir = "output/snv/mouse/tempdir",
        output_base_name = "mouse_snvs",
        ref_fasta = "/reference/mus_musculus/mm10/ucsc_assmebly/illumina_download/Sequence/WholeGenomeFasta/genome.fa",
        min_depth = c(5, 10, 20, 30),
        job_base = "mouse_sbatch",
        log_base = "slurmOut/mouse_snv",
        min_snvs_per_cluster = 200,
        n_bootstraps = 10000,
        max_prop_missing_at_site = 0.9,
        tree_image_type = "pdf",
        cleanup = TRUE,
        other_job_header_options = c(
            "--time=8:00:00",
            "--partition=himem,general"
        ),
        other_batch_options = c(
            "ml purge",
            "ml miniforge3",
            'eval "$(conda shell.bash hook)"'
        )
    )
```

## Prep the mouse samples with GFP/RFP
```{r scanbit_prep_mouse_gfp_rfp}
#| dependson: [read_raw_mouse_geo]
transgene_samples <-
    read_tsv(
        "misc/validation_geo_metadata.tsv",
        show_col_types = FALSE
    ) %>%
    filter(sample_group == "truthset validation") %>%
    pull(Sample_ID) %>%
    unique()

roberts_transgene_samples <-
    read_tsv(
        "misc/mouse_samples_Roberts_realign_metadata.tsv",
        show_col_types = FALSE
    ) %>%
    pull(Sample_ID)

ltbc_transgene_samples <- c("S0260", "S0261")

transgene_samples <- c(
    transgene_samples,
    roberts_transgene_samples,
    ltbc_transgene_samples
)

samples_worked <-
    parallel::mclapply(
        transgene_samples,
        mc.cores = 5,
        mc.preschedule = FALSE,
        function(this_sample) {
            scanbit_one_sample(
                this_sample = this_sample,
                species = "mouse",
                ploidy = "mm10",
                ref_fasta = "custom_ref/gfp_rfp_mm10/fasta/genome.fa",
                out_dir = "output/snv/transgene/"
            )
        }
    )
names(samples_worked) <- transgene_samples

print(samples_worked)
```

## Run scanBit on the human samples
Doing these individually
```{r scanbit_run_human}
#| dependson: read_raw_human
human_samples <-
    read_tsv(
        "misc/validation_geo_metadata.tsv",
        show_col_types = FALSE
    ) %>%
    filter(sample_group != "truthset validation") %>%
    pull(Sample_ID) %>%
    unique()

samples_worked <-
    parallel::mclapply(
        human_samples,
        mc.cores = 5,
        mc.preschedule = FALSE,
        function(this_sample) {
            scanbit_one_sample(
                this_sample = this_sample,
                species = "human",
                ploidy = "GRCh38",
                ref_fasta = "/home/gdrobertslab/lab/GenRef/10x-hg38/fasta/genome.fa",
                out_dir = "output/snv/human/"
            )
        }
    )
names(samples_worked) <- human_samples

print(samples_worked)
```
