## Annotate Cells

I realized that in order to call SNVs, I have to have cell types annotated. I'm going to do that now for the samples in which I've labelled tumor cells.

```{r}
if (!file.exists("output/seurat_objects/annotated_obs.qs")) {
    obs <- list.files("output/seurat_objects") %>%
        str_remove(".qs") %>%
        grep(pattern = "annotated", invert = TRUE, value = TRUE)

    ob_list <- list()
    for (ob in obs) {
        ob_list[[ob]] <- qs::qread(
            paste0("output/seurat_objects/", ob, ".qs")
        )
    }
    ob_list <- lapply(ob_list, JoinLayers)

    # annotate and save list of annotated objects for me to do stuff with tmr morning
    ob_list <- lapply(ob_list, annotate_celltypes, species = "mouse")
    qs::qsave(ob_list, "output/seurat_objects/annotated_obs.qs")
} else {
    ob_list <- qs::qread("output/seurat_objects/annotated_obs.qs")
}
```

### Visualize Annotations

```{r}
annot_plots <- list()
for (ob in names(ob_list)) {
    annot_plots[[ob]] <-
        dimplot_better(
            ob_list[[ob]],
            group = "cell_type"
        ) +
        NoLegend() |
            dimplot_better(
            ob_list[[ob]],
            group = "is_tumor"
        ) +
    ggtitle(label = ob)
}

wrap_plots(annot_plots)

# look at macrophage markers
# these markers are from panglao db
mac_markers <- c("Cd68", "Fcgr1", "Naaa", "March1")

mac_plots <- list()
for (ob in names(ob_list)) {
    mac_plots[[ob]] <- 
        r_feature_plot(
            ob_list[[ob]],
            features = mac_markers,
            coord.fixed = TRUE
        )
}

wrap_plots(mac_plots)
```

I feel confident that our tumor cells are actually tumor cells, since the cell type annotation labelled them as fibroblasts, which we commonly see happen in our single cell data. This cell type label matching the manual annotation based on the presence of our reporter plasmid confirms that these are our tumor cells and we can use these datasets as ground truth.

Additionally, the macrophage and monocyte populations are likely macrophages and monocytes, evidenced by the expression of canonical macrophage markers from panglaodb. These genes were also expressed in the cell types labelled monocytes and neutrophils, so our normal control cell types will be macrophages, monocytes, and neutrophils.

## Call SNVs

This is going to be following along with the vignette from kidcancerlab/scanBit repo.

```{r install-scanBit}
# devtools::install_github("kidcancerlab/scanBit")
bam_prefix <- list(
    "merged_GSE189856" = "counts/GSE189856/",
    "merged_GSE201615" = "counts/GSE201615/",
    "merged_GSE247228" = "counts/GSE247228/",
    "merged_S0260_61" = "/home/gdrobertslab/lab/Counts_2/"
)

bam_suffix <- list(
    "merged_GSE189856" = "/outs/possorted_genome_bam.bam",
    "merged_GSE201615" = "/outs/possorted_genome_bam.bam",
    "merged_GSE247228" = "/outs/possorted_genome_bam.bam",
    "merged_S0260_61" =  "/possorted_genome_bam.bam"
)

start <- Sys.time()
snp_tree <- list()
for (ob in names(ob_list)) {
    if (!dir.exists(paste0("output/", ob))) dir.create(paste0("output/", ob))
    # make table with columns cell_barcode, cell_group, and bam_file
    bc_bam_tbl <- 
        ob_list[[ob]]@meta.data %>%
        mutate(
            cell_group = paste(sample_id, seurat_clusters, sep = "_"),
            bam_file = paste0(bam_prefix[[ob]], ob_list[[ob]]$sample_id, bam_suffix[[ob]])
        ) %>%
        rownames_to_column("cell_barcode") %>%
        mutate(cell_barcode = str_remove(cell_barcode, "_[1-9]")) %>%
        select(cell_group, bam_file, cell_barcode) %>%
        as_tibble()
    
    # get snp tree
    snp_tree[[ob]] <-
        get_snp_tree(
            cellid_bam_table = bc_bam_tbl,
            temp_dir = paste0("output/", ob, "/temp"),
            ploidy = "mm10",
            output_dir = paste0("output/", ob),
            min_depth = c(5, 10, 20),
            output_base_name = ob,
            ref_fasta = "/home/gdrobertslab/lab/GenRef/10x-mm10/fasta/genome.fa",
            sbatch_base = "snv_call",
            other_sbatch_options = c(
                "--time=8:00:00"
            )
        )
}
end <- Sys.time()
end - start

```

