## Differential expression analyses
We are going to compare between each cell type between:

- PBS and Dox
- PBS and Magrolimab
- PBS and Dox+Magrolimab
- Dox and Dox+Magrolimab
- Magrolimab and Dox+Magrolimab

This will separate out the effect of each drug and the combination of both drugs.

### Make table of comparisons
```{r de_cmp_list}
#| cache.vars: [de_comps, p_adj_cutoff, fc_cutoff, min_de_genes]
de_comps <-
    list(
        pbs_vs_dox     = list(group_1 = "PBS", group_2 = "Dox"),
        pbs_vs_mag     = list(group_1 = "PBS", group_2 = "Magrolimab"),
        pbs_vs_dox_mag = list(group_1 = "PBS", group_2 = "Dox+Mag"),
        dox_vs_dox_mag = list(group_1 = "Dox", group_2 = "Dox+Mag"),
        mag_vs_dox_mag = list(group_1 = "Magrolimab", group_2 = "Dox+Mag")
    )

p_adj_cutoff <- 0.05
fc_cutoff <- 1
min_de_genes <- 50
```

### Perform differential expression analysis on mouse data
```{r de_analyses_mouse}
#| dependson: [merge_dimplots_mouse, cmp_list]
#| cache.vars: [de_res_list, fc_list]
merged_sobj_mouse <- qs::qread("output/seurat_objects/merged/merged_mouse.qs")

cell_types <- unique(merged_sobj_mouse$cell_type)
min_cells_group <- 20

de_res_list <- list()
fc_list <- list()

for (one_type in cell_types) {
    for (de_comparison in names(de_comps)) {
        label <- paste0(de_comparison, ".", one_type)
        cell_type_table <-
            merged_sobj_mouse@meta.data %>%
            filter(cell_type == one_type &
                   (Name == de_comps[[de_comparison]]$group_1 |
                    Name == de_comps[[de_comparison]]$group_2)) %>%
            group_by(Name) %>%
            summarise(n = n())

        # Only proceed if we have enough cells in both groups
        if (nrow(cell_type_table) == 2 &&
            min(cell_type_table$n, Inf) > min_cells_group) {
            de_res_list[[label]] <-
                subset(merged_sobj_mouse, cell_type == one_type) %>%
                FindMarkers(
                    group.by = "Name",
                    ident.1 = de_comps[[de_comparison]]$group_1,
                    ident.2 = de_comps[[de_comparison]]$group_2,
                ) %>%
                rownames_to_column("gene") %>%
                mutate(
                    significant = p_val_adj < p_adj_cutoff &
                    abs(avg_log2FC) > fc_cutoff
                )

            # Get fold change values for GSEA if we have some DE genes
            if (nrow(de_res_list[[label]]) > min_de_genes) {
                fc_list[[label]] <-
                    subset(merged_sobj_mouse, cell_type == one_type) %>%
                    FoldChange(
                        group.by = "Name",
                        ident.1 = de_comps[[de_comparison]]$group_1,
                        ident.2 = de_comps[[de_comparison]]$group_2
                    ) %>%
                    rownames_to_column("gene") %>%
                    arrange(desc(avg_log2FC)) %>%
                    pull(avg_log2FC, name = gene)
            }
        }
    }
}
```

### Differential analysis on human tumor cells
```{r de_analyses_human}
#| dependson: [merge_dimplots_human, cmp_list]
#| cache.vars: [de_res_list, fc_list]
merged_sobj_human <- qs::qread("output/seurat_objects/merged/merged_human.qs")

for (de_comparison in names(de_comps)) {
    de_res_list[[paste0(de_comparison, ".tumor")]] <-
        FindMarkers(
            merged_sobj_human,
            group.by = "Name",
            ident.1 = de_comps[[de_comparison]]$group_1,
            ident.2 = de_comps[[de_comparison]]$group_2,
        ) %>%
        rownames_to_column("gene") %>%
        mutate(
            significant = p_val_adj < p_adj_cutoff &
                abs(avg_log2FC) > fc_cutoff,
            group_1 = de_comps[[de_comparison]]$group_1,
            group_2 = de_comps[[de_comparison]]$group_2
        ) %>%
        left_join(
            subset(
                merged_sobj_human,
                Name == de_comps[[de_comparison]]$group_1 |
                    Name == de_comps[[de_comparison]]$group_2
            ) %>%
                AverageExpression(group.by = "Name") %>%
                unlist() %>%
                as.data.frame() %>%
                rownames_to_column("gene") %>%
                rename_with(
                    ~ str_remove(., "^RNA.") %>% paste0("_avg_expr"),
                    .cols = -gene
                )
        )

    n_de_genes <-
        filter(
            de_res_list[[paste0(de_comparison, ".tumor")]],
            significant == TRUE
        ) %>%
        nrow()

    # Get fold change values for GSEA if we have some DE genes
    if (n_de_genes > min_de_genes) {
        fc_list[[paste0(de_comparison, ".tumor")]] <-
            FoldChange(
                merged_sobj_human,
                group.by = "Name",
                ident.1 = de_comps[[de_comparison]]$group_1,
                ident.2 = de_comps[[de_comparison]]$group_2
            ) %>%
            rownames_to_column("gene") %>%
            arrange(desc(avg_log2FC)) %>%
            pull(avg_log2FC, name = gene)
    }
}
```

## Create volcano plots and save DE results
```{r de_volcano_plots}
#| results: "asis"
#| dependson: [de_analyses_mouse, de_analyses_human]
volc_plots <- list()
for (comparison in names(de_res_list)) {
    if (nrow(de_res_list[[comparison]]) > 0) {
        volc_plots[[comparison]] <-
            de_res_list[[comparison]] %>%
            ggplot(
                aes(
                    x = avg_log2FC,
                    y = -log10(p_val),
                    color = significant
                )
            ) +
            geom_point(alpha = 0.5) +
            ggrepel::geom_text_repel(
                data = . %>% filter(significant),
                aes(label = gene),
                size = 2
            ) +
            labs(
                x = "log2 fold change",
                y = "-log10(p-value)",
                title = comparison,
                color = "Significant"
            )
        ggsave(
            paste0("output/figures/volcano_plots/volcano_", comparison, ".png"),
            volc_plots[[comparison]],
            width = 6,
            height = 5
        )

        de_res_list[[comparison]] %>%
            write_tsv(paste0("output/degs/", comparison, ".tsv"))
    }
}

# Make a tabset of volcano plots
qreport::maketabs(volc_plots)
```

## Venn diagram of tumor DE genes for different comparisons
This will show how many genes are differentially expressed across the different comparisons.
```{r de_venn_diagram}
#| dependson: de_analyses_human
#| fig.height: 10
#| fig.width: 10
tumor_comps <-
    lapply(
        de_res_list[grep("tumor", names(de_res_list), value = TRUE)],
        function(x) {
            filter(x, significant == TRUE) %>%
                pull(gene)
    })

tumor_venn <-
    ggvenn::ggvenn(tumor_comps, padding = 0.5)
ggsave(
    "output/figures/venn_tumor_degs.pdf",
    tumor_venn,
    width = 10,
    height = 10
)
tumor_venn
```

## GSEA analysis
This will identify pathways that are enriched in the differentially expressed genes.
```{r de_gsea}
#| results: "asis"
#| dependson: [de_analyses_human, de_analyses_mouse, helper_gsea_plot]

#GSEA Miltilevel
cat_tib <- dplyr::tribble(
    ~category, ~subcategory,   ~cat_expl,
    "H",        "",            "Hallmark_paths",
    "C2",      "CP:KEGG",      "KEGG",
    "C3",      "TFT:GTRD",     "Transcription_factor_targets",
    "C5",      "GO:BP",        "GO Biological Process",
    "C5",      "GO:MF",        "GO Molecular Function",
    "C6",       "",            "Oncogenic_signature")

gsea_plot_list <- list()

set.seed(1337)
for (cmp_group in names(fc_list)) {
    species <-
        if_else(
            grepl("tumor", cmp_group),
            "Homo sapiens",
            "Mus musculus"
        )
    for (i in seq_len(nrow(cat_tib))) {
        gsea_ref <-
            msigdbr::msigdbr(
                species = species,
                category = cat_tib$category[i],
                subcategory = cat_tib$subcategory[i]
            ) %>%
                split(x = .$gene_symbol, f = .$gs_name)

        gsea_output <-
            fgsea::fgseaMultilevel(
                gsea_ref,
                fc_list[[cmp_group]],
                minSize = 10,
                maxSize = 500,
                nPermSimple = 10000
            ) %>%
                filter(padj < 0.05) %>%
                arrange(desc(NES)) %>%
                select(-leadingEdge) %>%
                mutate(
                    order = seq_len(n()),
                    justify_y = if_else(NES > 0, 1, 0),
                    y_pos = if_else(NES > 0, -0.1, 0.1)
                )

        if (nrow(gsea_output) > 0) {
            write_tsv(
                gsea_output,
                paste0("output/gsea/gsea_", cmp_group, ".tsv")
            )

            gsea_plot_list[[paste(cmp_group, cat_tib$cat_expl[i])]] <-
                gsea_dotplot(gsea_output, x_col = "NES")

            ggsave(
                paste0(
                    "output/figures/gsea/gsea_",
                    cmp_group, "_",
                    cat_tib$cat_expl[i],
                    ".png"
                ),
                gsea_plot_list[[paste(cmp_group, cat_tib$cat_expl[i])]],
                width = 6,
                height = 5
            )
        }

    }
}

qreport::maketabs(gsea_plot_list)
```
