
```{r}
merged_sobj_mouse_ctrls <-
    qs::qread("output/seurat_objects/merged/merged_mouse.qs")

# Grab just data from our control mice
merged_sobj_mouse_ctrls <-
    merged_sobj_mouse_ctrls[
        ,
        merged_sobj_mouse_ctrls$Sample_ID %in% c("S0066", "S0074")
    ]

merged_sobj_mouse_ctrls$cell_group <-
    paste0(
        merged_sobj_mouse_ctrls$Sample_ID,
        "_c",
        merged_sobj_mouse_ctrls$individ_clusters
    )

min_n_cells <- 500

keep_clusters <-
    merged_sobj_mouse_ctrls$cell_group %>%
    table() %>%
    as.data.frame() %>%
    filter(Freq > min_n_cells) %>%
    pull(".")

# Keep only clusters with > 900 cells
merged_sobj_mouse_ctrls <-
    merged_sobj_mouse_ctrls %>%
    subset(cell_group %in% keep_clusters)
```


## Loop through downsample cutoffs and run scanBit
```{r mouse_scanbit}
ncell_cutoffs <- c(1, 2, 3, 4, 5, 10, 20, 30, 50, 100, 200, 500)

Idents(merged_sobj_mouse_ctrls) <-
    merged_sobj_mouse_ctrls$cell_group

output <-
    parallel::mclapply(
        rev(ncell_cutoffs),
        mc.cores = 10,
        mc.preschedule = FALSE,
        function(cutoff) {
            set.seed(1337)
            # Get the cell barcodes for the current cutoff
            cell_barcode_group_table <-
                merged_sobj_mouse_ctrls %>%
                subset(downsample = cutoff) %>%
                FetchData(vars = c(
                    "cell_group",
                    "cell_barcode",
                    "bam_file",
                    "nCount_RNA"
                )) %>%
                as_tibble()

            write_tsv(
                cell_barcode_group_table,
                file.path(
                    "output/downsample/cell_tables",
                    paste0("cell_barcode_table_", cutoff, ".txt")
                )
            )

            cell_barcode_group_table <-
                select(cell_barcode_group_table, -nCount_RNA)

            dir.create(
                paste0(
                    "output/downsample/scanbit_out/",
                    "downsample_",
                    cutoff,
                    "/tempdir"
                ),
                recursive = TRUE,
                showWarnings = FALSE
            )

            snp_tree <-
                scanBit::get_snp_tree(
                    cellid_bam_table = cell_barcode_group_table,
                    ploidy = "mm10",
                    temp_dir = paste0(
                        "output/downsample/scanbit_out/",
                        "downsample_",
                        cutoff,
                        "/tempdir"
                    ),
                    output_dir = paste0(
                        "output/downsample/scanbit_out/",
                        "downsample_",
                        cutoff
                    ),
                    output_base_name = paste0("downsample_", cutoff),
                    ref_fasta = "/reference/mus_musculus/mm10/ucsc_assmebly/illumina_download/Sequence/WholeGenomeFasta/genome.fa",
                    min_depth = 5,
                    sbatch_base = "downsample",
                    min_snvs_per_cluster = 1,
                    n_bootstraps = 10000,
                    max_prop_missing_at_site = 0.9,
                    cleanup = TRUE,
                    other_sbatch_options = "--time=24:00:00"
                )
        }
    )
```


## Try with human Sample_ID
```{r downsample_human1}
x_14 <-
    tenx_load_qc(
        "/home/gdrobertslab/lab/Counts_2/X00014/filtered_feature_bc_matrix",
        violin_plot = FALSE
    ) %>%
    subset(nCount_RNA <= 30000 & percent.mt < 20) %>%
    process_seurat()

x_14$bam_file <-
    "/home/gdrobertslab/lab/Counts_2/X00014/possorted_genome_bam.bam"

x_14$cell_barcode <- Cells(x_14)
opt_res <-
    rrrSingleCellUtils::optimize_silhouette(x_14, summary_plot = FALSE) %>%
    dplyr::arrange(dplyr::desc(sil_vals)) %>%
    dplyr::pull(res_vals) %>%
    head(1)

x_14 <- Seurat::FindClusters(x_14, resolution = opt_res)

# source("/home/gdrobertslab/lab/Analysis/Matt/dev/rrrSingleCellUtils/R/recurl.R")

# x_14 <-
#     recurluster(
#         x_14,
#         max_recursions = 3,
#         parallel = TRUE,
#         verbose = FALSE,
#         do_plots = FALSE
#     )

DimPlot(
    x_14,
    group.by = c("RNA_snn_res.0.25"),
    label = TRUE,
    label.box = TRUE,
    label.size = 2,
    repel = TRUE
) +
    NoLegend() +
    FeaturePlot(
        x_14,
        features = "COL1A1"
    )

x_14$cell_group <- paste0("clust_", x_14$RNA_snn_res.0.25)

min_n_cells <- 200

keep_clusters <-
    x_14$cell_group %>%
    table() %>%
    as.data.frame() %>%
    filter(Freq > min_n_cells) %>%
    pull(".")

# Keep only clusters with > 900 cells
x_14 <-
    x_14 %>%
    subset(cell_group %in% keep_clusters)

Idents(x_14) <-
    x_14$cell_group
```


```{r downsample_x14_umi}
n_umi_cutoffs <- c(1, 2.5, 5, 25, 50, 100, 200, 400) * 10000


output <-
    parallel::mclapply(
        rev(n_umi_cutoffs),
        mc.cores = 10,
        mc.preschedule = FALSE,
        function(cutoff) {
            out_dir <-
                paste0(
                    "output/downsample/scanbit_out/x14/downsample_",
                    sprintf("%.0f", cutoff),
                    "_umi"
                )

            dir.create(
                file.path(out_dir, "tempdir"),
                recursive = TRUE,
                showWarnings = FALSE
            )

            # Get the cell barcodes for the current cutoff
            cell_barcode_group_table <- make_cbt(x_14, max_umi = cutoff)

            write_tsv(
                cell_barcode_group_table,
                file.path(
                    out_dir,
                    paste0(
                        "cell_barcode_table_x14_",
                        sprintf("%.0f", cutoff),
                        "_umi.txt"
                    )
                )
            )

            cell_barcode_group_table <-
                select(cell_barcode_group_table, -nCount_RNA)

            snp_tree <-
                scanBit::get_snp_tree(
                    cellid_bam_table = cell_barcode_group_table,
                    ploidy = "GRCh38",
                    temp_dir = file.path(out_dir, "tempdir"),
                    output_dir = out_dir,
                    output_base_name = paste0(
                        "downsample_x14_",
                        sprintf("%.0f", cutoff)
                    ),
                    ref_fasta = "/home/gdrobertslab/lab/GenRef/10x-hg38/fasta/genome.fa",
                    min_depth = 5,
                    sbatch_base = "downsample_x14",
                    min_snvs_per_cluster = 10,
                    n_bootstraps = 10000,
                    max_prop_missing_at_site = 0.9,
                    cleanup = TRUE,
                    other_sbatch_options = "--time=24:00:00"
                )
        }
    )
```

## Downsample by cell number instead of umi
```{downsample_x14_cells}
n_cell_cutoffs <- c(1, 2, 3, 4, 5, 10, 20, 50, 100, 250, 500)

output <-
    parallel::mclapply(
        n_cell_cutoffs,
        mc.cores = 10,
        mc.preschedule = FALSE,
        function(cutoff) {
            out_dir <-
                paste0(
                    "output/downsample/scanbit_out/x14/",
                    "downsample_",
                    sprintf("%.0f", cutoff),
                    "_cells"
                )

            dir.create(
                file.path(out_dir, "tempdir"),
                recursive = TRUE,
                showWarnings = FALSE
            )

            # Get the cell barcodes for the current cutoff
            cell_barcode_group_table <- make_cbt(x_14, n_cells = cutoff)

            write_tsv(
                cell_barcode_group_table,
                file.path(
                    out_dir,
                    paste0(
                        "cell_barcode_table_x14_",
                        sprintf("%.0f", cutoff),
                        "_cells.txt"
                    )
                )
            )

            cell_barcode_group_table <-
                select(cell_barcode_group_table, -nCount_RNA)

            snp_tree <-
                scanBit::get_snp_tree(
                    cellid_bam_table = cell_barcode_group_table,
                    ploidy = "GRCh38",
                    temp_dir = file.path(out_dir, "tempdir"),
                    output_dir = out_dir,
                    output_base_name = paste0(
                        "downsample_x14_",
                        sprintf("%.0f", cutoff)
                    ),
                    ref_fasta = "/home/gdrobertslab/lab/GenRef/10x-hg38/fasta/genome.fa",
                    min_depth = 5,
                    sbatch_base = "downsample_x14_cells",
                    min_snvs_per_cluster = 10,
                    n_bootstraps = 10000,
                    max_prop_missing_at_site = 0.9,
                    cleanup = TRUE,
                    other_sbatch_options = "--time=24:00:00"
                )
        }
    )
```



## Loop over all the osteo datasets and run scanBit
```{r downsample_loop_x}
samples_tsv_x <-
    read_tsv("misc/sample_metadata.tsv", show_col_types = FALSE) %>%
    filter(grepl("X00", Sample_ID))

min_n_cells <- 200
min_n_cells <- 1

n_cell_cutoffs <- c(1, 2, 3, 4, 5, 10, 20, 50, 100, 100000, 500000)

output <-
    parallel::mclapply(
        samples_tsv_x$Sample_ID,
        mc.cores = 5,
        mc.preschedule = FALSE,
        function(this_sample) {
            sobj <-
                qs::qread(paste0(
                    "output/seurat_objects/individual/",
                    this_sample,
                    "_human.qs"
                ))

            sobj$cell_group <- paste0("clust_", sobj$individ_clusters)

            keep_clusters <-
                sobj$cell_group %>%
                table() %>%
                as.data.frame() %>%
                filter(Freq > min_n_cells) %>%
                pull(".")

            # Keep only clusters with > min_n_cells cells
            sobj <-
                sobj %>%
                subset(cell_group %in% keep_clusters)

            Idents(sobj) <-
                sobj$cell_group

            scanbit_downsample_loop(
                sobj,
                cutoff_vector = n_cell_cutoffs,
                out_dir_base = file.path(
                    "output/downsample/scanbit_out/x_samples/",
                    this_sample
                ),
                sample_label = this_sample
            )
        }
    )
```


stuff  <- qs::qread("output/seurat_objects/individual/X00021_human.qs")

stuff_cnv <-
    SCEVAN::pipelineCNA(
        GetAssayData(stuff, layer = "counts"),
        par_cores = 5,
        norm_cell = WhichCells(stuff, idents = 4),
        ClonalCN = TRUE
    )

    plotCNA_withAnnotCells



```{r testplot}
genotype_code <-
    c(
        #"./." = 10,
        "0/0" = 1,
        "0/1" = 2,
        "1/1" = 3
    )
#stuff <-
    read_tsv("test.vcf") %>%
    rename(chr = "#CHROM") %>%
    select(-ID, -QUAL, -FILTER, -INFO, -FORMAT) %>%
    pivot_longer(
        cols = starts_with("clust_"),
        names_to = "cluster",
        values_to = "call"
    ) %>%
    separate_wider_delim(
        call,
        delim = ":",
        names = c("GT", "PL", "DP", "AD", "GQ")
    ) %>%
    select(-PL, -AD, -GQ) %>%
    filter(GT %in% c("0/0", "0/1", "1/1")) %>%
    mutate(genotype_num = genotype_code[GT]) %>%
    mutate(chr_pos = paste(chr, POS)) %>%
    select(chr_pos, cluster, genotype_num) %>%
    pivot_wider(
        id_cols = chr_pos,
        names_from = "cluster",
        values_from = genotype_num
    ) %>%
    column_to_rownames("chr_pos") %>%
    arrange(clust_0, clust_1, clust_2, clust_3) %>%
    as.matrix() %>%
    pheatmap::pheatmap(cluster_cols = FALSE, cluster_rows = FALSE)


    #mutate(depth = str_split(DP, ",") %>% as.vector() %>% sum())
    ggplot(aes(x = cluster, y = chr_pos, fill = GT)) +
    geom_tile()

```

```{r}
bootstrap_downsample_plot <-
    readxl::read_xlsx("downsample_bootstrap_values.xlsx") %>%
    mutate(
        cells_char = as.factor(cells_char) %>% fct_reorder(cells),
    ) %>%
    ggplot(
        aes(
            x = cells_char,
            y = bootstrap_val,
            group = sample,
            color = sample
        )
    ) +
    geom_boxplot(aes(group = cells), color = "black", fill = "lightgray") +
    #geom_point(size = 2, alpha = 0.5) +
    #geom_line() +
    scale_color_brewer(palette = "Set2") +
    labs(
        title = "Top Level Bootstrap Values for Downsampled Data",
        subtitle = "Confidence When Using Only 'X' Cells",
        x = "Cells per cluster",
        y = "Bootstrap value of top tree division"
    )

ggsave(
    "output/figures/downsample_bootstrap_plot.png",
    width = 8,
    height = 6
)
```

