## Here I am going to downsample the mouse samples to varying number of cells per cluster to run scanBit
This is to test how low we can go while the algorithm still works and provide bcf files to use to quantify coverage per cell in later analyses
```{r}
merged_sobj_mouse_ctrls <-
    qs::qread("output/seurat_objects/merged/merged_norm_mouse.qs")

min_n_cells <- 500

keep_clusters <-
    merged_sobj_mouse_ctrls$cell_group %>%
    table() %>%
    as.data.frame() %>%
    filter(Freq > min_n_cells) %>%
    pull(".")

# Keep only clusters with > min_n_cells cells
merged_sobj_mouse_ctrls <-
    merged_sobj_mouse_ctrls %>%
    subset(cell_group %in% keep_clusters)

# Here, 500000 is forcing us to keep all the cells, so functions as a control
n_cell_cutoffs <- c(1, 2, 3, 4, 5, 10, 20, 50, 100, 500, 500000)

output <- scanbit_downsample_loop(
    merged_sobj_mouse_ctrls,
    cutoff_vector = n_cell_cutoffs,
    out_dir_base = "output/downsample/scanbit_out/mouse/",
    sample_label = "mouse",
    min_read_depth = 5,
    ploidy = "mm10",
    ref_fasta = "/home/gdrobertslab/lab/GenRef/10x-mm10/fasta/genome.fa"
)
```

## Loop over all the GEO human datasets and run scanBit
```{r downsample_loop_geo}
#| dependson: [helper_downsample_scanbit, read_raw_human_geo]
samples_geo <-
    read_tsv("misc/validation_geo_metadata.tsv", show_col_types = FALSE) %>%
    filter(grepl("tumor validation|healthy samples", sample_group)) %>%
    pull(Sample_ID) %>%
    unique()

min_n_cells <- 200

n_cell_cutoffs <- c(1, 2, 3, 4, 5, 10, 20, 50, 100, 500, 500000)

output <-
    parallel::mclapply(
        samples_geo,
        mc.cores = 5,
        mc.preschedule = FALSE,
        function(this_sample) {
            sobj <-
                qs::qread(paste0(
                    "output/seurat_objects/individual/",
                    this_sample,
                    "_human.qs"
                ))

            sobj$cell_group <- paste0("clust_", sobj$individ_clusters)

            keep_clusters <-
                sobj$cell_group %>%
                table() %>%
                as.data.frame() %>%
                filter(Freq > min_n_cells) %>%
                pull(".")

            # Keep only clusters with > min_n_cells cells
            sobj <-
                sobj %>%
                subset(cell_group %in% keep_clusters)

            Idents(sobj) <-
                sobj$cell_group

            scanbit_downsample_loop(
                sobj,
                cutoff_vector = n_cell_cutoffs,
                out_dir_base = file.path(
                    "output/downsample/scanbit_out/geo_samples/",
                    this_sample
                ),
                sample_label = this_sample,
                ploidy = "GRCh38",
                ref_fasta = "/home/gdrobertslab/lab/GenRef/10x-hg38/fasta/genome.fa"
            )
        }
    )
```

## Show how bootstrap values degrade with decreasing cell count and UMI per pseudobulked cluster
The data were gathered by hand since I don't have a programatic way to pull the top node bootstrap values out of the pdfs
```{r downsample_plot}
rna_per_run <-
    list.files(
        path = "output/downsample/scanbit_out",
        pattern = "cell_barcode_table*",
        recursive = TRUE,
        full.names = TRUE
    ) %>%
    read_tsv(id = "run") %>%
    select(-cell_barcode, -bam_file) %>%
    mutate(
        run = basename(run) %>%
            str_remove("cell_barcode_table_") %>%
            str_remove("_cells.txt")
    ) %>%
    separate_wider_delim(
        run,
        delim = "_",
        names = c("sample", "cells")
    ) %>%
    mutate(cells = as.numeric(cells)) %>%
    group_by(sample, cells, cell_group) %>%
    summarize(total_umi = sum(nCount_RNA), .groups = "drop") %>%
    group_by(sample, cells) %>%
    summarize(mean_umi = mean(total_umi), .groups = "drop")

bootstrap_downsample_data <-
    readxl::read_xlsx("downsample_bootstrap_values.xlsx") %>%
    full_join(rna_per_run) %>%
    left_join(
        bind_rows(
            read_tsv(
                "misc/validation_geo_metadata.tsv",
                show_col_types = FALSE
            ) %>%
                select(Sample_ID, sample_group, tumor_type) %>%
                rename(sample = Sample_ID) %>%
                distinct(),
            tibble(sample = "mouse", sample_group = "Mouse")
        )
    ) %>%
    mutate(
        cells_char = as.factor(cells_char) %>% fct_reorder(cells),
    ) %>%
    filter(cells > 1 & sample_group != "Mouse") #!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!! Do separate plot for mouse!!!!!!!!!!!!!

bootstrap_ncell_plot <-
    bootstrap_downsample_data %>%
    ggplot(
        aes(
            x = cells_char,
            y = bootstrap_val,
            group = sample
        )
    ) +
    geom_boxplot(aes(group = cells), color = "black", fill = "lightgray") +
    ggbeeswarm::geom_beeswarm(aes(color = tumor_type), size = 2) +
    facet_wrap(~ sample_group, ncol = 1) +
    labs(
        title = "Top Level Bootstrap Values for Downsampled Data",
        subtitle = "Confidence When Using Only 'X' Cells",
        x = "Cells per cluster",
        y = "Bootstrap value of top tree division"
    )

ggsave(
    "output/figures/downsample_bootstrap_cells_plot.pdf",
    plot = bootstrap_ncell_plot,
    width = 8,
    height = 10
)

bootstrap_numi_plot <-
    bootstrap_downsample_data %>%
    ggplot(
        aes(
            x = mean_umi,
            y = bootstrap_val,
            group = sample
        )
    ) +
    geom_point(aes(color = tumor_type), size = 2) +
    geom_vline(xintercept = 250000, color = "red", linetype = "dashed") +
    facet_wrap(~ sample_group, ncol = 1) +
    scale_x_continuous(
        breaks = scales::breaks_width(5000000)
    ) +
    labs(
        title = "Top Level Bootstrap Values for Downsampled Data",
        subtitle = "Confidence When Using Only 'X' UMIs per Cluster. Vertical line at 250,000",
        x = "Mean UMI per cluster",
        y = "Bootstrap value of top tree division",
        color = "Sample type"
    ) +
    theme(
        legend.position = "bottom"
    )

ggsave(
    "output/figures/downsample_bootstrap_UMI_plot.pdf",
    plot = bootstrap_numi_plot,
    width = 10,
    height = 8
)
```
