## Read metadata, read data and make prepreocessed seurat objects
Load in Seurat objects and save individual qs objects off to disk

## Read in F420 and K7M2 samples to use as SingleR references
```{r read_raw_mouse_references}
#| dependson: [helper_read_raw_fun]
read_raw_from_table(
    sample_table = read_tsv(
        "misc/mouse_samples_Roberts_metadata.tsv",
        show_col_types = FALSE
    ) %>%
        filter(Tissue == "cell culture"),
    h5_folder = "/home/gdrobertslab/lab/Counts_2/",
    organism = "mouse",
    annotate_species = "mouse",
    count_table_file = "output/counts/mouse_culture_cell_counts.txt",
    bam_file = "possorted_genome_bam.bam"
)
```

### Table of cell culture mouse cell counts per sample
```{r read_raw_mouse_culture_cell_count_table}
#| dependson: read_raw_mouse_references
DT::datatable(read_tsv("output/counts/mouse_culture_cell_counts.txt"))
```

## Prep known tumor cells to add to SingleR ref
```{r read_raw_prep_mouse_ref}
#| dependson: [helper_read_raw_fun]
culture_ref_samples <-
    read_tsv(
        "misc/mouse_samples_Roberts_metadata.tsv",
        show_col_types = FALSE
    ) %>%
    filter(Tissue == "cell culture") %>%
    pull(Sample_ID)

tumor_refs <-
    lapply(
        culture_ref_samples,
        function(x) {
            set.seed(1337)
            sobj <- qs::qread(paste0(
                "output/seurat_objects/individual/",
                x,
                "_mouse.qs"
            )) %>%
                subset(downsample = 500)

            return(
                list(
                    sce = as.SingleCellExperiment(sobj),
                    labels = rep("Tumor", ncol(sobj))
                )
            )
        }
    )

qs::qsave(tumor_refs, "output/refs/mouse_tumor_refs.qs")
```

## Read in B6, Balbc, F420 and K7M2 samples
```{r read_raw_mouse}
#| dependson: [read_raw_prep_mouse_ref]
tumor_refs <- qs::qread("output/refs/mouse_tumor_refs.qs")

read_raw_from_table(
    sample_table = read_tsv(
        "misc/mouse_samples_Roberts_metadata.tsv",
        show_col_types = FALSE
    ) %>%
        filter(Tissue != "cell culture"),
    h5_folder = "/home/gdrobertslab/lab/Counts_2/",
    organism = "mouse",
    annotate_species = "mouse",
    count_table_file = "output/counts/mouse_cell_counts.txt",
    bam_file = "possorted_genome_bam.bam",
    annotate_additional_ref = lapply(
        tumor_refs,
        function(x) x$sce
    ),
    annotate_additional_labels = lapply(
        tumor_refs,
        function(x) x$labels
    )
)
```

### Table of mouse cell counts per sample
```{r read_raw_mouse_cell_count_table}
#| dependson: read_raw_mouse
DT::datatable(read_tsv("output/counts/mouse_cell_counts.txt"))
```

## Read in Robert's lab data that was re-aligned
```{r read_raw_realigned}
tumor_refs <- qs::qread("output/refs/mouse_tumor_refs.qs")

read_raw_from_table(
    sample_table = read_tsv(
        "misc/mouse_samples_Roberts_realign_metadata.tsv",
        show_col_types = FALSE
    ),
    h5_folder = "output/cellranger_out",
    organism = "mouse",
    annotate_species = "mouse",
    count_table_file = "output/counts/mouse_cell_counts.txt",
    bam_file = "possorted_genome_bam.bam",
    annotate_additional_ref = lapply(
        tumor_refs,
        function(x) x$sce
    ),
    annotate_additional_labels = lapply(
        tumor_refs,
        function(x) x$labels
    )
)
```

### Table of mouse cell counts per sample
```{r read_raw_mouse_realigned_cell_count_table}
#| dependson: read_raw_realigned
DT::datatable(read_tsv("output/counts/mouse_cell_counts.txt"))
```

## Mouse flex data from 10x
```{r read_raw_flex}
#| dependson: [helper_read_raw_fun, get_data_flex]
read_raw_from_table(
    sample_table = read_tsv(
        "misc/flex_metadata.tsv",
        show_col_types = FALSE
    ),
    h5_folder = "input/10x_flex",
    organism = "mouse",
    annotate_species = "mouse",
    count_table_file = "output/counts/mouse_cell_counts_flex.txt",
    bam_file = "count_sample_alignments.bam"
)
```

### Table of 10x flex cell counts per sample
```{r read_raw_flex_cell_count_table}
#| dependson: read_raw_flex
DT::datatable(read_tsv("output/counts/mouse_cell_counts_flex.txt"))
```

## Make Seurat objects from human GEO data
```{r read_raw_human_geo}
#| dependson: [helper_read_raw_fun, get_data_count_geo]
read_raw_from_table(
    sample_table = read_tsv(
        "misc/validation_geo_metadata.tsv",
        show_col_types = FALSE
    ) %>%
        filter(
            sample_group == "tumor validation" |
                sample_group == "healthy samples"
        ) %>%
        select(-sra_number, -fake_lane) %>%
        distinct(),
    h5_folder = "output/cellranger_out",
    organism = "human",
    annotate_species = "human",
    count_table_file = "output/counts/human_geo_cell_counts.txt",
    bam_file = "possorted_genome_bam.bam"
)
```

## Table of GEO human cell counts per sample
```{r read_raw_human_geo_count_table}
#| dependson: read_raw_human_geo
DT::datatable(read_tsv("output/counts/human_geo_cell_counts.txt"))
```

## Get data from GEO mouse data
```{r read_raw_mouse_geo}
#| dependson: [helper_read_raw_fun, get_data_count_geo]
read_raw_from_table(
    sample_table = read_tsv(
        "misc/validation_geo_metadata.tsv",
        show_col_types = FALSE
    ) %>%
        filter(sample_group == "truthset validation") %>%
        select(-sra_number, -fake_lane) %>%
        distinct(),
    h5_folder = "output/cellranger_out",
    organism = "mouse",
    annotate_species = "mouse",
    count_table_file = "output/counts/mouse_geo_cell_counts_geo.txt",
    bam_file = "possorted_genome_bam.bam"
)
```

### Table of GEO mouse data
```{r read_raw_mouse_geo_count_table}
#| dependson: read_raw_mouse_geo
DT::datatable(read_tsv("output/counts/mouse_geo_cell_counts_geo.txt"))
```
