Before I download all the available data, I'm going to test it on the control samples from GSE247228.

The sra Run numbers are as follows:
    - SRR26705626
    - SRR26705627
    - SRR26705629

## Get FASTQ Files

I'll be using SRA-Toolkit/2.10.8 for this. Specifically, the `fasterq-dump` function.

### GSE247228

```{bash get-computing-power}
srun_l -c 12 -p himem --x11 --pty bash
```

```{bash get-fastqs}
ml SRA-Toolkit/2.10.8
fasterq-dump --outdir input/fastqs/GSE247228 --progress SRR26705626 SRR26705627 SRR26705629 SRR26705628 SRR26705630 SRR26705631
```

### GSE179077

Fluorescent UnaG in hypoxic glioblastoma cells

```{bash get-unag-fastqs}
mkdir input/fastqs/GSE179077
fasterq-dump --outdir input/fastqs/GSE179077 --progress SRR29428580 SRR29428581 SRR29428582 SRR29428583 SRR29428584 SRR29428585 SRR29428586 SRR29428587
```

### GSE201615

Does it have dsRED?

For this I'm thinking I make the new reference containing dsRED, cas9... and maybe Ccne-dsRED conjugate? I can't tell exactly what's going on with their experimental methods.


```{bash}
#trying parallel-fastq-dump
conda env create parallel_fastq_dump.yml
conda activate parallel_fastq_dump
parallel-fastq-dump --outdir input/fastqs/GSE201615/parallel -t 3 --split-files -s SRR18932563 SRR18932566 SRR18932567


fastq-dump --outdir input/fastqs/GSE201615/tmp --split-files SRR18932563 SRR18932566 SRR18932567

#we have 3 fastqs for each; one is the Index and the other two are our Read files
#_1 is I1
#_2 is R1
#_3 is R2


cd input/fastqs/GSE201615
for sra in SRR18932562 SRR18932563 SRR18932566 SRR18932567; do
    mkdir $sra
    mv ${sra}_1.fastq \
        ${sra}/${sra}_S1_I1_001.fastq
    mv ${sra}_2.fastq \
        ${sra}/${sra}_S1_R1_001.fastq
    mv ${sra}_3.fastq \
        ${sra}/${sra}_S1_R2_001.fastq
done

cd ../../..
```

## Update Reference Genome

I need to add the fluorescent reporter sequence to our reference genome before aligning our fastqs. I'll be adding it to our 10x-mm10 reference genome.

First I'm going to make a copy of the mm10 gtf and fastq.

```{bash copy-fastq-gtf}
cp /home/gdrobertslab/lab/GenRef/10x-mm10/fasta/genome.fa custom_ref/mm10_genome.fa
cp /home/gdrobertslab/lab/GenRef/10x-mm10/genes/genes.gtf.gz custom_ref/mm10_genes.gtf.gz
chmod 744 custom_ref/
```

### eGFP

I will first need to create a file named GFP.fa containing the nucleotide sequence for our GFP protein.

```{bash make-new-ref}
#get number of bases in GFP.fa
grep -v "^>" custom_ref/eGFP/eGFP.fa | tr -d "\n" | wc
#720 bases

#lets make a custom gtf for gfp; \t helps us make new columns that are needed in the gtf; we need both the exon and transcript lines 
echo -e 'eGFP\tunknown\texon\t1\t720\t.\t+\t.\tgene_id "eGFP"; transcript_id "eGFP"; gene_biotype "protein_coding";\n'\
'eGFP\tunknown\ttranscript\t1\t720\t.\t+\t.\tgene_id "eGFP"; transcript_id "eGFP"; gene_biotype "protein_coding";\n'\
'eGFP\tunknown\tgene\t1\t720\t.\t+\t.\tgene_id "eGFP"; gene_name "eGFP"; gene_biotype "protein_coding";' > custom_ref/eGFP/eGFP.gtf

#copy mm10 reference FASTA over (so we can modify it and not screw up the og)

#append GFP.fa to end of mm10 genome
cp custom_ref/mm10_genome.fa custom_ref/eGFP/genome.fa
cat custom_ref/eGFP/eGFP.fa >> custom_ref/eGFP/genome.fa

#check that GFP "chromosome" is present
grep ">" custom_ref/eGFP/genome.fa

#copy over original gtf and gunzip
cp custom_ref/mm10_genes.gtf custom_ref/eGFP/genes.gtf.gz
gunzip custom_ref/eGFP/genes.gtf.gz

#append GFP gtf to mm10 gene file
cat custom_ref/eGFP/eGFP.gtf >> custom_ref/eGFP/genes.gtf

#Time to make our new reference using cellranger mkref
cellranger mkref --output-dir custom_ref/eGFP/demo_eGFP \
    --genome demo_eGFP \
    --fasta custom_ref/eGFP/genome.fa \
    --genes custom_ref/eGFP/genes.gtf
```

### HRE-UnaG

Now I'm going to make a genome containing the hypoxic reporter element plus unag.

```{bash make-unag-ref}
#making a folder within custom ref for this, otherwise it'll get too cluttered
mkdir custom_ref/unag/

grep -v "^>" custom_ref/unag/hre_unag.fa | tr -d "\n" | wc
#8646 bases


echo -e 'HRE-dUnaG\tunknown\texon\t1\t8646\t.\t+\t.\tgene_id "HRE-dUnaG"; transcript_id "HRE-dUnaG"; gene_biotype "protein_coding";\n'\
'HRE-dUnaG\tunknown\ttranscript\t1\t8646\t.\t+\t.\tgene_id "HRE-dUnaG"; transcript_id "HRE-dUnaG"; gene_biotype "protein_coding";\n'\
'HRE-dUnaG\tunknown\tgene\t1\t8646\t.\t+\t.\tgene_id "HRE-dUnaG"; gene_name "HRE-dUnaG"; gene_biotype "protein_coding";' > custom_ref/unag/hre_unag.gtf

#Going to append the gtf and fasta to the mm10
#append GFP.fa to end of mm10 genome
cp custom_ref/mm10_genome.fa custom_ref/unag/genome.fa

cp custom_ref/mm10_genes.gtf.gz custom_ref/unag/genes.gtf.gz
gunzip custom_ref/unag/genes.gtf.gz

#cat gtf and fasta for hre-unag to mm10 reference files
cat custom_ref/unag/hre_unag.fa >> custom_ref/unag/genome.fa
cat custom_ref/unag/hre_unag.gtf >> custom_ref/unag/genes.gtf

#make reference
ml cellranger
cellranger mkref --output-dir custom_ref/unag/unag_ref \
    --genome unag_ref \
    --fasta custom_ref/unag/genome.fa \
    --genes custom_ref/unag/genes.gtf \
    --jobmode=local \
    --localcores=8
```

### RFP

also known as dsRed

I'm going to make a reference with just RFP and see if downsampled fastqs have any reads that map to it.

```{bash make-rfp-ref}
mkdir custom_ref/RFP
grep -v "^>" custom_ref/RFP/RFP.fa | tr -d "\n" | wc
#672 bases

echo -e 'RFP\tunknown\texon\t1\t672\t.\t+\t.\tgene_id "RFP"; transcript_id "RFP"; gene_biotype "protein_coding";\n'\
'RFP\tunknown\ttranscript\t1\t672\t.\t+\t.\tgene_id "RFP"; transcript_id "RFP"; gene_biotype "protein_coding";\n'\
'RFP\tunknown\tgene\t1\t672\t.\t+\t.\tgene_id "RFP"; gene_name "RFP"; gene_biotype "protein_coding";' > custom_ref/RFP/RFP.gtf

#copy over fasta and gtf files, and gunzip gtf
cp custom_ref/mm10_genome.fa custom_ref/RFP/genome.fa
cp custom_ref/mm10_genes.gtf.gz custom_ref/RFP/genes.gtf.gz
gunzip custom_ref/RFP/genes.gtf.gz

#give self write permission
chmod 774 custom_ref/RFP/*

#cat RFP gtf and fasta to mm10 files
cat custom_ref/RFP/RFP.fa >> custom_ref/RFP/genome.fa
cat custom_ref/RFP/RFP.gtf >> custom_ref/RFP/genes.gtf

#make just RFP reference
ml cellranger
cellranger mkref --output-dir custom_ref/RFP/full_RFP \
    --genome full_RFP \
    --fasta custom_ref/RFP/genome.fa \
    --genes custom_ref/RFP/genes.gtf \
    --jobmode=local
```

## Run cellranger count

### On eGFP data From GSE247228

```{bash mk-bams}
# We need to convert our fastq names to a cellranger compatible format

for sra in SRR26705626 SRR26705627 SRR26705628 SRR26705629 SRR26705630 SRR26705631
do
    mkdir input/fastqs/GSE247228/${sra}
    mv input/fastqs/GSE247228/${sra}_1.fastq \
        input/fastqs/GSE247228/${sra}/${sra}_S1_R1_001.fastq
    mv input/fastqs/GSE247228/${sra}_2.fastq \
        input/fastqs/GSE247228/${sra}/${sra}_S1_R2_001.fastq
done

#run cellranger count!
sbatch cellranger_count_egfp.sh
```

### On unag data from GSE179077

```{bash mk-bams-GSE179077}
#rename to cellranger compatible format
for sra in SRR29428580 SRR29428581 SRR29428582 SRR29428583 SRR29428584 SRR29428585 SRR29428586 SRR29428587
do
    mkdir input/fastqs/GSE179077/${sra}
    mv input/fastqs/GSE179077/${sra}_1.fastq \
        input/fastqs/GSE179077/${sra}/${sra}_S1_R1_001.fastq
    mv input/fastqs/GSE179077/${sra}_2.fastq \
        input/fastqs/GSE179077/${sra}/${sra}_S1_R2_001.fastq
done

#run cellranger count
sbatch cellranger_count_unag.sh
```

### On GSE201615

```{bash GSE201615-cellranger-count}
sbatch cellranger_count_rfp.sh
```

## Look at Counts for Our Reporter

### GSE247228

```{r}
#cutoffs for feature hists
cutoff_tbl <- tribble(      ~sample,     ~feature, ~min_val, ~max_val,
                      "SRR26705626", "nCount_RNA",     1000,    40000,
                      "SRR26705626", "percent.mt",        0,       10, 
                      "SRR26705627", "nCount_RNA",     1000,    60000,
                      "SRR26705627", "percent.mt",        0,       10,
                      "SRR26705628", "nCount_RNA",     1000,    50000,
                      "SRR26705628", "percent.mt",        0,       10,
                      "SRR26705629", "nCount_RNA",     1000,    75000,
                      "SRR26705629", "percent.mt",        0,       10,
                      "SRR26705630", "nCount_RNA",     1000,    50000,
                      "SRR26705630", "percent.mt",        0,       10,
                      "SRR26705631", "nCount_RNA",     1000,    40000, 
                      "SRR26705631", "percent.mt",        0,       10) 

#read in data
sample_names <- grep("SRR", list.files("counts/GSE247228/"), value = TRUE)
obj_list <- lapply(sample_names, function(x) {
    sobj <- tenx_load_qc(paste0("counts/GSE247228/",
                                x,
                                "/outs/filtered_feature_bc_matrix"),
                         violin_plot = FALSE)
    sobj$eGFP_exp <- GetAssayData(sobj, layer = "counts")["eGFP", ]
    subset(sobj, features = rownames(sobj)[rownames(sobj) != "eGFP"])
    # sobj@assays$RNA$counts <- sobj@assays$RNA$counts[rownames(sobj) != "eGFP", ]
    return(sobj)
})
names(obj_list) <- sample_names

qc_plts <- lapply(sample_names, function(x) {
    cuts <- subset(cutoff_tbl, sample == x) %>%
        select(feature, min_val, max_val)
    ob <- obj_list[[x]]
    feature_hist(ob,
                 features = c("nCount_RNA", "percent.mt"),
                 cutoff_table = cuts) +
        ggtitle(x)
})

png("output/figures/GSE247228/qc_grid.png", height = 2000, width = 1600)
grid.arrange(grobs = qc_plts)
dev.off()

#qc and make egfp feature plots
obj_list <- lapply(sample_names, function(x) {
    ob <- obj_list[[x]]
    cuts <- subset(cutoff_tbl, sample == x) %>%
        column_to_rownames("feature")
    ob <- subset(ob,
                 nCount_RNA > cuts["nCount_RNA", "min_val"] &
                 nCount_RNA < cuts["nCount_RNA", "max_val"] &
                 percent.mt < cuts["percent.mt", "max_val"]) %>%
        process_seurat()
    ob$sample_id <- x
    png(paste0("output/figures/GSE247228/dimplot_egfp_", x, ".png"),
        width = 1000,
        height = 600)
    print(dimplot_better(ob, group = "seurat_clusters") |
    r_feature_plot(ob, features = "eGFP_exp"))
    dev.off()
    
    return(ob)
})

merged_obs <- merge(obj_list[[1]], obj_list[-1]) %>%
    process_seurat()
merged_obs <- JoinLayers(merged_obs)
#check for batch effects
dimplot_better(merged_obs, "sample_id")

#Going to binarize eGFP_exp
merged_obs$binarized_eGFP <- merged_obs$eGFP_exp > 0

dimplot_better(merged_obs, group  = "seurat_clusters")| r_feature_plot(merged_obs, features = "binarized_eGFP", pt.size = 0.3)
```

It seems like clusters 6, 9, and 11 are our eGFP expressing cells. I'm going to save this off as a seurat object as well as just a metadata file.

```{r write-off-GSE247228}
merged_obs$is_tumor <- merged_obs$seurat_clusters %in% c(6, 9, 11)

qs::qsave(merged_obs, "output/seurat_objects/merged_GSE247288.qs")
rownames_to_column(merged_obs@meta.data, "bc") %>%
    write.table(file = "output/metadata/metadata_GSE247288.tsv",
                sep = "\t",
                row.names = FALSE,
                quote = FALSE)
```

### GSE179077

```{r unag-exp}
sample_names <- grep("SRR", list.files("counts/GSE179077"), value = TRUE)
obj_list <- lapply(sample_names, function(x) {
    ob <- tenx_load_qc(paste0("counts/GSE179077/",
                        x,
                        "/outs/filtered_feature_bc_matrix"),
                 violin_plot = FALSE)
    ob$unag_exp <- GetAssayData(ob, layer = "counts")["HRE-dUnaG", ]
    
    png(paste0("output/figures/GSE179077/qc_", x, ".png"),
        height = 600,
        width = 800)
    feature_hist(ob,
                 features = c("nCount_RNA", "percent.mt"),
                 cutoff_table = tribble(    ~feature, ~min_val, ~max_val,
                                        "nCount_RNA",     1000,    10000,
                                        "percent.mt",        0,       10))
    dev.off()
    
    ob <- subset(ob, features = rownames(ob)[rownames(ob) != "HRE-dUnaG"],
                     subset = nCount_RNA >= 1000 &
                     nCount_RNA <= 10000 &
                     percent.mt < 10) %>%
        process_seurat()
    return(ob)
})
names(obj_list) <- sample_names

merged_gbm <- merge(obj_list[[1]], obj_list[-2]) %>%
    process_seurat()

#binarize unag_exp
merged_gbm$unag_present <- merged_gbm$unag_exp > 0

r_feature_plot(merged_gbm, features = c("unag_present", "Sox2", "S100a4", "Cd4", "Cd8"), pt.size = 0.3, coord.fixed = TRUE)
```