## Run scanBit on individual cells with min_read_depth of 1 to use for read depth and accuracy Analysis
For this I'm keeping ten individual cells per cluster, and each cell is assigned to its own cell_group so that I am calculating read depth data for many individual cells.
```{r readdepth_scanbit}
samples_geo <-
    read_tsv("misc/validation_geo_metadata.tsv", show_col_types = FALSE) %>%
    filter(grepl("tumor validation|healthy samples", sample_group)) %>%
    pull(Sample_ID) %>%
    unique()

n_cell_cutoffs <- 1

# Keeping mc.cores low since each run submits 100+ jobs
output <-
    parallel::mclapply(
        samples_geo,
        mc.cores = 2,
        mc.preschedule = FALSE,
        function(this_sample) {
            sobj <-
                qs::qread(paste0(
                    "output/seurat_objects/individual/",
                    this_sample,
                    "_human.qs"
                ))

            sobj$cell_group <-
                paste0(
                    "clust_",
                    sobj$individ_clusters,
                    "_",
                    sobj$cell_barcode
                )

            Idents(sobj) <- sobj$individ_clusters
            set.seed(1337)
            sobj <- subset(sobj, downsample = 10)

            Idents(sobj) <-
                sobj$cell_group

            scanbit_downsample_loop(
                sobj,
                cutoff_vector = n_cell_cutoffs,
                min_read_depth = c(1, 2, 5, 10, 20),
                out_dir_base = file.path(
                    "output/read_depth/scanbit_out/geo_samples/",
                    this_sample
                ),
                keep_temp_dir = TRUE,
                sample_label = this_sample,
                ploidy = "GRCh38",
                ref_fasta = "/home/gdrobertslab/lab/GenRef/10x-hg38/fasta/genome.fa"
            )
        }
    )
```

## Batch job to count site coverage
```{bash readdepth_count_depth}
sbatch sbatchCmds/geo_bcf_depths.sh
```

## Calculate the # of positions covered where any cell has a variant
```{r readdepth_var_positions}
#| cache.vars: [var_depths, ave_var_depths]
all_var_files <-
    list.files(
        path = "output/read_depth/depths/bcf_depth",
        pattern = "variant_pos",
        full.names = TRUE
    )

var_depths <-
    read_tsv(
        all_var_files,
        id = "cell_sample",
        show_col_types = FALSE,
        skip_empty_rows = FALSE
    ) %>%
    mutate(
        cell_sample = basename(cell_sample) %>%
            str_remove("variant_pos_") %>%
            str_remove(".txt"),
        group = "Variant sites"
    )

ave_var_depths <-
    var_depths %>%
    pivot_wider(
        names_from = "depth",
        values_from = "count",
        values_fill = 0
    ) %>%
    pivot_longer(
        cols = c(-cell_sample, -group),
        names_to = "depth",
        values_to = "count"
    ) %>%
    group_by(depth) %>%
    summarize(
        mean_count_var = mean(count),
        se_count = sd(count) / sqrt(n()),
        .groups = "drop",
    ) %>%
    mutate(depth = as.numeric(depth)) %>%
    arrange(depth) %>%
    slice_head(n = 30)
```

## Compile all single-cell variant call depths
```{r readdepth_compile_depths}
#| cache.vars: [all_depths, ave_depths]
all_depth_files <-
    list.files(
        path = "output/read_depth/depths/bcf_depth",
        pattern = "bcf_depth_G",
        full.names = TRUE
    )
all_depths <-
    read_tsv(
        all_depth_files,
        id = "cell_sample",
        show_col_types = FALSE,
    ) %>%
    mutate(
        cell_sample = basename(cell_sample) %>%
            str_remove("bcf_depth_") %>%
            str_remove(".txt"),
        group = "All sites"
    )

ave_depths <-
    all_depths %>%
    pivot_wider(
        names_from = "depth",
        values_from = "count",
        values_fill = 0
    ) %>%
    pivot_longer(
        cols = c(-cell_sample, -group),
        names_to = "depth",
        values_to = "count"
    ) %>%
    group_by(depth) %>%
    summarize(
        mean_count = mean(count),
        se_count = sd(count) / sqrt(n()),
        .groups = "drop"
    ) %>%
    mutate(depth = as.numeric(depth)) %>%
    arrange(depth) %>%
    slice_head(n = 50)
```

```{r readdepth_plot_coverage}
site_covered_plot <-
    ggplot() +
    geom_point(
        data = ave_depths,
        aes(x = depth, y = mean_count),
        color = "black"
    ) +
    geom_point(
        data = ave_var_depths,
        aes(x = depth, y = mean_count_var),
        color = "red"
    ) +
    scale_y_log10(labels = scales::label_comma()) +
    labs(
        title = "Mean Sites Covered to a Given Depth",
        subtitle = "In Variant Call Data on Single Cells",
        x = "Depth",
        y = "Average number of positions covered"
    )

ggsave(
    "output/figures/mean_coverage_1_cell.png",
    site_covered_plot,
    width = 6,
    height = 4
)

jitter_coverage_plot <-
    rbind(all_depths, var_depths) %>%
    filter(depth <= 20) %>%
    pivot_wider(
        names_from = "depth",
        values_from = "count",
        values_fill = 0
    ) %>%
    pivot_longer(
        cols = c(-cell_sample, -group),
        names_to = "depth",
        values_to = "count"
    ) %>%
    mutate(
        depth_factor = as.factor(depth) %>%
            fct_reorder(as.numeric(depth))
    ) %>%
    ggplot(
        aes(
            x = depth_factor,
            y = count,
            color = group
        )
    ) +
    geom_jitter(height = 0, alpha = 0.5, size = 0.5) +
    scale_y_log10(labels = scales::label_comma()) +
    scale_color_manual(values = plot_cols[1:2]) +
    labs(
        title = "Number of Total and Non-reference Sites Covered by Single Cells at Varying Depths",
        x = "Read depth of variant call",
        y = "Number of sites covered at each depth",
        color = NULL
    ) +
    theme(
        legend.position = c(0.9, 0.9)
    )

ggsave(
    "output/figures/total_and_nonref_site_coverage.pdf",
    jitter_coverage_plot,
    width = 8,
    height = 6
)

rm(all_depths, var_depths, ave_depths, ave_var_depths)
```

## Calculate expected number of mistakes for each depth
Based on flex accuracy calculations
```{r readdepth_accuracy_calc}
#| cache.vars: [expected_mis_calls]
# Make a named vector or accuracy - name is depth, value is % accurate
flex_accuracy_prop <-
    read_tsv("output/flex/flex_accuracy.txt", show_col_types = FALSE) %>%
    pull(mean_perc_wrong, name = depth) / 100

expected_mis_calls <-
    ave_depths %>%
    mutate(exp_bad_variants = mean_count * flex_accuracy_prop[depth])

expected_errors_plot <-
    expected_mis_calls %>%
    head(n = 100) %>%
    ggplot() +
    geom_point(aes(x = depth, y = mean_count), color = "black") +
    geom_point(aes(x = depth, y = exp_bad_variants), color = "red", size = 2) +
    scale_y_log10(labels = scales::label_comma())

ggsave(
    "output/figures/expected_errors_1_cell.png",
    expected_errors_plot,
    width = 6,
    height = 4
)
```

```{bash readdepth_actual_n_vars}
sbatch sbatchCmds/count_single_cell_variant_calls.sh
```

# Calculate the average number of variants actually called by depth
```{r readdepth_compile_depths_ave}
all_call_files <-
    list.files(
        path = "output/read_depth/depths/call_count",
        pattern = "bcf_calls_depth_G0",
        full.names = TRUE
    )
all_call_depths <-
    read_tsv(
        all_call_files,
        id = "cell_sample",
        show_col_types = FALSE,
        col_names = c("chr", "alt", "depth")
    ) %>%
    mutate(
        cell_sample = basename(cell_sample) %>%
            str_remove("bcf_calls_depth_") %>%
            str_remove(".txt")
    )

average_num_calls <-
    all_call_depths %>%
    group_by(depth, cell_sample) %>%
    summarize(n_vars = n(), .groups = "drop") %>%
    group_by(depth) %>%
    summarize(mean_n_vars = mean(n_vars))

n_vars_plot <-
    average_num_calls %>%
    head(n = 50) %>%
    ggplot() +
    geom_point(aes(x = depth, y = mean_n_vars), color = "black") +
    labs(
        title = "Average Number of Variants Called by Depth",
        subtitle = "In Variant Call Data on Single Cells",
        x = "Depth",
        y = "Mean number of variants called per cell"
    )

ggsave(
    "output/figures/average_vars_1_cell.png",
    n_vars_plot,
    width = 6,
    height = 4
)

calls_table <-
    full_join(expected_mis_calls, average_num_calls, by = "depth")


calls_summary_plot <-
    calls_table %>%
    rename(wrong_variants = exp_bad_variants, called_variants = mean_n_vars) %>%
    pivot_longer(
        cols = c(wrong_variants, called_variants),
        names_to = "var_type",
        values_to = "count"
    ) %>%
    mutate(
        var_type = str_replace(var_type, "_", " ") %>%
            str_to_title()
    ) %>%
    filter(depth <= 50) %>%
    ggplot(aes(x = depth, y = count, group = depth)) +
    geom_line() +
    geom_point(aes(color = var_type, shape = var_type), size = 2) +
    scale_color_manual(values = c("black", "red"), name = NULL) +
    scale_shape_manual(values = c(19, 4), name = NULL) +
    labs(
        title = "Total and Inaccurate Variant Calls",
        x = "Depth",
        y = "Number of variant calls"
    ) +
    theme(
        legend.position = c(0.85, 0.9),
        legend.background = element_blank()
    )

ggsave(
    "output/figures/wrong_vs_called_nvars_1_cell.png",
    calls_summary_plot,
    width = 6,
    height = 4
)

percent_wrong_plot <-
    calls_table %>%
    filter(depth <= 50) %>%
    rename(wrong_variants = exp_bad_variants, called_variants = mean_n_vars) %>%
    mutate(perc_wrong = wrong_variants / called_variants) %>%
    ggplot(
        aes(
            x = depth,
            y = perc_wrong
        )
    ) +
    geom_point() +
    scale_y_continuous(labels = scales::label_percent()) +
    labs(
        title = "Percent of Inaccurate Variant Calls by Depth",
        x = "Depth",
        y = "Percent inaccurate variant calls"
    )

ggsave(
    "output/figures/percent_wrong_1_cell.png",
    percent_wrong_plot,
    width = 6,
    height = 4
)
```

## Calculate how many sites have shared variant information per sample
```{bash readdepth_calc_shared}
#| dependson: [readdepth_scanbit]
sbatch sbatchCmds/count_single_cell_shared_sites.sh
```

```{r readdepth_shared_sites}
#| dependson: [readdepth_calc_shared]
shared_files <-
    list.files(
        path = "output/read_depth/shared",
        pattern = "shared.+\\.txt",
        full.names = TRUE
    )

parse_shared <- function(file_name) {
    dataset <- read_tsv(
        file_name,
        name_repair = "minimal",
        show_col_types = FALSE,
        id = "dataset"
    ) |>
        mutate(
            dataset = str_remove(dataset, ".+shared_") |>
                str_remove(".txt")#,
            #cell_1 = str_replace(cell_1, "-", ".")
        ) |>
        pivot_longer(
            cols = starts_with("clust"),
            names_to = "cell_2",
            values_to = "shared_sites"
        ) |>
        filter(cell_1 != cell_2) |>
        separate(dataset, into = c("sample", "cutoff"), sep = "_") %>%
        mutate(cutoff = factor(cutoff, levels = c("1", "2", "5", "10", "20")))

    return(dataset)
}

shared_sites <-
    lapply(
        shared_files,
        parse_shared
    ) %>%
    bind_rows()


shared_sites %>%
    filter(sample == "G0001" & cutoff == 1) %>%
    select(cell_1, cell_2, shared_sites) %>%
    pivot_wider(
        names_from = cell_2,
        values_from = shared_sites,
        values_fill = NA
    ) %>%
    column_to_rownames("cell_1") %>%
    pheatmap::pheatmap(
        silent = TRUE,
        fontsize = 5,
        filename = "output/figures/shared_sites_G0001_c1.pdf",
        width = 10,
        height = 10
    )

shared_sites_histograms <-
    shared_sites %>%
    ggplot(
        aes(
            x = shared_sites,
            fill = cutoff
        )
    ) +
    geom_histogram(bins = 200) +
    scale_x_continuous(n.breaks = 20) +
    facet_wrap(
        ~ cutoff,
        ncol = 1,
        scales = "free_y"
    ) +
    labs(
        title = "Distribution of Number of Genomic Positions With Shared Variant\nInformation Between Two Cells",
        subtitle = "For sites where any cell has a non-reference variant call",
        x = "Shared covered genomic positions",
        y = "Count",
        fill = "Minimum read depth\nto call variants"
    )

ggsave(
    "output/figures/shared_sites_histograms.pdf",
    plot = shared_sites_histograms,
    width = 8,
    height = 8
)

shared_sites %>%
    group_by(cutoff) %>%
    summarize(
        mean_shared_sites = mean(shared_sites),
        median_shared_sites = median(shared_sites)
    ) %>%
    write_tsv("output/read_depth/shared_sites_stats.tsv")
```
