## Test snp calling on flex data

### Download 10x flex data
```{bash download_flex}
#| eval: false

sbatch sbatchCmds/get10x_flex.sh
```

### Read in 10x flex data
```{r read_flex}
#| cache.vars: sobj_list
merged_sobj_mouse_flex <-
    qs::qread("output/seurat_objects/merged/merged_mouse.qs")

# Grab just data from our lab - excludes flex data
merged_sobj_mouse_flex <-
    merged_sobj_mouse_flex[, grepl("^flex_", merged_sobj_mouse_flex$Sample_ID)] %>%
    process_seurat()

merged_sobj_mouse_flex$cell_group <-
    paste0(
        merged_sobj_mouse_flex$Sample_ID,
        "_c",
        merged_sobj_mouse_flex$individ_clusters
    )

flex_dimplot <-
    (DimPlot(
        merged_sobj_mouse_flex,
        group.by = "Name",
        label = TRUE,
        label.box = TRUE,
        label.size = 2,
        repel = TRUE,
        ncol = 1
    ) +
        NoLegend()) +
    (DimPlot(
        merged_sobj_mouse_flex,
        group.by = "cell_group",
        label = TRUE,
        label.box = TRUE,
        label.size = 2,
        repel = TRUE,
        ncol = 1
    ) +
        NoLegend()) +
    (DimPlot(
        merged_sobj_mouse_flex,
        group.by = "cell_type",
        label = TRUE,
        label.box = TRUE,
        label.size = 2,
        repel = TRUE,
        ncol = 1
    ) +
        NoLegend())

ggsave(
    "output/figures/dimplot_flex_cell_type_group.pdf",
    flex_dimplot,
    width = 10,
    height = 20
)
```


```{r runSNVs}
cell_barcode_group_table <-
    merged_sobj_mouse_flex@meta.data %>%
    select(cell_group, cell_barcode, bam_file) %>%
    as_tibble()

snp_tree <-
    get_snp_tree(
        cellid_bam_table = cell_barcode_group_table,
        ploidy = "mm10",
        output_dir = "output/snv/flex/",
        temp_dir = "output/snv/flex/tempdir",
        output_base_name = "flex_snvs",
        ref_fasta = "/reference/mus_musculus/mm10/ucsc_assmebly/illumina_download/Sequence/WholeGenomeFasta/genome.fa",
        min_depth = c(1, 5, 10, 20, 30),
        sbatch_base = "flex_sbatch",
        slurm_base = "slurmOut/flex_snv",
        min_snvs_per_cluster = 500,
        n_bootstraps = 10000,
        max_prop_missing_at_site = 0.9,
        cleanup = TRUE,
        other_sbatch_options = "--time=8:00:00"
    )

# This was erroring out because there aren't enough snvs
# I lowered min_snvs_per_cluster to 50 to get it to work

scanBit::group_clusters_by_dist(
    merged_bcf_file = "output/snv/flex/mergedflex_snvs_c5.bcf",
    output_file = "output/snv/flex/mergedflex_snvs_c5_min_5.tsv",
    min_snvs_per_cluster = 5,
    max_prop_missing_at_site = 0.99,
    tree_figure_file = "output/snv/flex/mergedflex_snvs_c5_tree.png",
    other_sbatch_options = "",
    temp_dir = "."
)
```

## Quantify the number of SNVs for each level of coverage

Ultimately want a plot of SNVs vs read depth

```{bash flex_cov_alts}
sbatch

```

```{r flex_prop_wrong}
all_flex_files <-
    list.files(
        path = "output/flex/var_counts_d1",
        pattern = ".txt",
        full.names = TRUE
    )

# In the future, break this down by mt vs non-mt to see if the non-mt are the weird inaccurate calls ~ 500 - 1000 depth
flex_accuracy <-
    read_tsv(
        all_flex_files,
        id = "sample",
        show_col_types = FALSE,
        col_names = c("chr", "alt", "depth", "count")
    ) %>%
    mutate(
        sample = basename(sample) %>%
            str_remove(".txt"),
        alt = if_else(
            alt == ".",
            "ref",
            "variant"
        )
    ) %>%
    group_by(sample, depth, alt) %>%
    summarize(count = sum(count, na.rm = TRUE), .groups = "drop") %>%
    group_by(sample, depth) %>%
    mutate(total_count = sum(count, na.rm = TRUE)) %>%
    ungroup() %>%
    pivot_wider(
        id_cols = c(sample, depth, total_count),
        names_from = "alt",
        values_from = "count",
        values_fill = 0
    ) %>%
    mutate(perc_wrong = variant / total_count * 100) %>%
    group_by(depth) %>%
    summarize(
        mean_total_count = mean(total_count, na.rm = TRUE),
        mean_perc_wrong = mean(perc_wrong, na.rm = TRUE),
        .groups = "drop"
    ) %>%
    filter(depth != 1)

write_tsv(
    flex_accuracy,
    "output/flex/flex_accuracy.txt"
)

accuracy_plot <-
    filter(flex_accuracy, depth <= 50) %>%
    ggplot(aes(x = depth, y = mean_perc_wrong / 100)) +
    #geom_smooth(alpha = 0.5) +
    geom_point() +
    geom_text(
        data = filter(flex_accuracy, depth <= 10),
        aes(label = depth),
        nudge_y = 0.00001
    ) +
    scale_y_continuous(labels = scales::label_percent()) +
    labs(
        title = "Percent Inaccurate Calls by Read Depth",
        x = "Read depth of variant calls",
        y = "% inaccurate variant calls"
    )

ggsave(
    "output/figures/flex_accuracy_scatterplot.png",
    plot = accuracy_plot,
    width = 6,
    height = 4
)
```

## Calculate how many positions covered to each depth by normal data
```{r flex_actual_covered}
all_flex_files <-
    list.files(
        path = "output/flex/var_counts_d1",
        pattern = ".txt",
        full.names = TRUE
    )

# In the future, break this down by mt vs non-mt to see if the non-mt are the weird inaccurate calls ~ 500 - 1000 depth
flex_accuracy <-
    read_tsv(
        all_flex_files,
        id = "sample",
        show_col_types = FALSE,
        col_names = c("chr", "alt", "depth", "count")
    ) %>%
    mutate(
        sample = basename(sample) %>%
            str_remove(".txt"),
        alt = if_else(
            alt == ".",
            "ref",
            "variant"
        )
    ) %>%
    group_by(sample, depth, alt) %>%
    summarize(count = sum(count, na.rm = TRUE), .groups = "drop") %>%
    group_by(sample, depth) %>%
    mutate(total_count = sum(count, na.rm = TRUE)) %>%
    ungroup() %>%
    pivot_wider(
        id_cols = c(sample, depth, total_count),
        names_from = "alt",
        values_from = "count",
        values_fill = 0
    ) %>%
    mutate(perc_wrong = variant / ref * 100) %>%
    group_by(depth) %>%
    summarize(
        mean_total_count = mean(total_count, na.rm = TRUE),
        mean_perc_wrong = mean(perc_wrong, na.rm = TRUE),
        .groups = "drop"
    ) %>%
    filter(depth != 1)
```

## Plot how many sites are covered at each depth

## Plot how many inaccurate variant calls we would expect at each depth


read_tsv("output/snv/mouse/alt_pos_counts.txt")


ggplot(stuff, aes(x = depth, y = sites_cov)) +
    geom_point() +
    scale_y_log10()

ggplot(stuff, aes(x = depth, y = wrong)) +
    geom_point() +
    scale_y_log10() +
    scale_x_log10() +
    coord_fixed() +
    geom_smooth(show.legend = TRUE)

ggplot(stuff, aes(x = depth, y = wrong / sites_cov)) +
    geom_point() +
    #scale_y_log10() +
    scale_x_log10()
#!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!! Can I split this up by Chr? maybe keep mt chr and toss other chrs data where depth > 100

head(stuff, n = 1000) %>%
    ggplot(aes(x = depth, y = cumul_wrong)) +
    geom_point() +
    scale_y_log10() +
    scale_x_log10()

head(stuff, n = 1000) %>%
    ggplot(aes(x = depth, y = cumul_wrong / cumul_sites * 100)) +
    geom_point() +
    scale_y_log10() +
    scale_x_log10()

```
