Before I download all the available data, I'm going to test it on the control samples from GSE247228.

The sra Run numbers are as follows:
    - SRR26705626
    - SRR26705627
    - SRR26705629

## Get FASTQ Files

I'll be using SRA-Toolkit/2.10.8 for this. Specifically, the `fasterq-dump` function.

```{bash get-computing-power}
srun_l -c 12 -p himem --x11 --pty bash
```

```{bash get-fastqs}
ml SRA-Toolkit/2.10.8
fasterq-dump --outdir input/fastqs --progress SRR26705626 SRR26705627 SRR26705629
```

## Update Reference Genome

I need to add the GFP sequence to our reference genome before aligning our fastqs. I'll be adding it to our 10x-mm10 reference genome. Our new reference will be in custom_ref/10x-mm10-gfp

I will first need to create a file named GFP.fa containing the nucleotide sequence for our GFP protein.

```{bash make-new-ref}
#get number of bases in GFP.fa
grep -v "^>" custom_ref/GFP.fa | tr -d "\n" | wc
#922 bases

#lets make a custom gtf for gfp; \t helps us make new columns that are needed in the gtf; we need both the exon and transcript lines 
echo -e 'GFP\tunknown\texon\t1\t922\t.\t+\t.\tgene_id "GFP"; transcript_id "GFP"; gene_biotype "proetin_coding";\n'\
'GFP\tunknown\ttranscript\t1\t922\t.\t+\t.\tgene_id "GFP"; transcript_id "GFP"; gene_biotype "proetin_coding";\n'\
'GFP\tunknown\tgene\t1\t922\t.\t+\t.\tgene_id "GFP"; gene_name "GFP"; gene_biotype "proetin_coding";' > custom_ref/GFP.gtf

#copy mm10 reference FASTA over (so we can modify it and not screw up the og)
cp /home/gdrobertslab/lab/GenRef/10x-mm10/fasta/genome.fa custom_ref/genome.fa
chmod 744 custom_ref/*

#append GFP.fa to end of mm10 genome
cat custom_ref/GFP.fa >> custom_ref/genome.fa

#check that GFP "chromosome" is present
grep ">" custom_ref/genome.fa

#copy over original gtf and gunzip
cp /home/gdrobertslab/lab/GenRef/10x-mm10/genes/genes.gtf.gz custom_ref/.
chmod 744 custom_ref/*
gunzip custom_ref/genes.gtf.gz

#append GFP gtf to mm10 gene file
cat custom_ref/GFP.gtf >> custom_ref/genes.gtf

#Time to make our new reference using cellranger mkref
cellranger mkref --output-dir custom_ref/demo_GFP \
    --genome demo_GFP \
    --fasta custom_ref/genome.fa \
    --genes custom_ref/genes.gtf
```

## See why no GFP reads

### Gonna try aligning to reference with just gfp
```{bash just-gfp-ref}
cellranger mkref --output-dir custom_ref/just_GFP \
    --genome just_GFP \
    --fasta custom_ref/GFP.fa \
    --genes custom_ref/GFP.gtf

mkdir just_gfp_counts
cellranger count \
    --id SRR26705626 \
    --output-dir just_gfp_counts/SRR26705626 \
    --fastqs input/fastqs/SRR26705626/ \
    --localcores 10 \
    --transcriptome custom_ref/just_GFP \
    --nosecondary \
    --disable-ui \
    --create-bam true

## see how many reads map to gfp chromosome
samtools idxstat just_gfp_counts/SRR26705626/outs/possorted_genome_bam.bam
```

### Going to add gfp sequences to a fastq and try aligning that 

```{bash just-gfp-ref}
#Going to append my gfp sequence to my fastq
#downsample junk_fastq
mkdir -p small_fastqs/out
ml seqtk
mkdir -p small_fastqs/SRR26705626
seqtk sample -s100 input/fastqs/SRR26705626/SRR26705626_S1_R1_001.fastq 10000 > small_fastqs/SRR26705626/SRR26705626_S1_R1_001.fastq
seqtk sample -s100 input/fastqs/SRR26705626/SRR26705626_S1_R2_001.fastq 10000 > small_fastqs/SRR26705626/SRR26705626_S1_R2_001.fastq

# I then manually added in 50 gfp-derived reads into fastq R1, replacing the last 50 reads

#Going to call cellranger count with GFP only reference
ml cellranger
cellranger count \
    --id SRR26705626 \
    --output-dir small_fastqs/out/SRR26705626 \
    --fastqs small_fastqs/SRR26705626 \
    --localcores 10 \
    --transcriptome custom_ref/just_GFP \
    --nosecondary \
    --disable-ui \
    --create-bam true

#see how many reads map to gfp chromosome
samtools idxstat small_fastqs/out/SRR26705626/outs/possorted_genome_bam.bam
```


## Make BAMs Aligned to New Genome

```{r mk-bams}
# We need to convert our fastq names to a cellranger compatible format
# I'll just be doing this for SRR26
# mv input/fastqs/SRR26705626/SRR26705626_1.fastq \
#     input/fastqs/SRR26705626/SRR26705626_S1_R1_001.fastq
# mv input/fastqs/SRR26705626/SRR26705626_2.fastq \
#     input/fastqs/SRR26705626/SRR26705626_S1_R2_001.fastq

for sra in SRR26705626 SRR26705627 SRR26705629
do
    mv input/fastqs/${sra}/${sra}_1.fastq \
        input/fastqs/${sra}/${sra}_S1_R1_001.fastq
    mv input/fastqs/${sra}/${sra}_2.fastq \
        input/fastqs/${sra}/${sra}_S1_R2_001.fastq
done

#run cellranger count!
sbatch cellranger_count.sh
```

## Look at New Object

```{r}
sobj <- tenx_load_qc("counts/SRR26705626/outs/filtered_feature_bc_matrix", violin_plot = FALSE)

sum(rownames(sobj) == "GFP")

feature_hist(sobj,
             features = c("nCount_RNA", "percent.mt"),
             cutoff_table =
    tribble(~feature, ~min_val, ~max_val,
            "nCount_RNA", 1200, 50000,
            "percent.mt", 0, 10))

sobj <- subset(sobj,
               nCount_RNA >= 1200 &
               nCount_RNA <= 50000 &
               percent.mt < 10) %>%
    process_seurat()

r_feature_plot(sobj, features = c("GFP", "Gfp", "gfp"))
```

Weirdly, I'm not seeing any GFP expression in any of the cells. I'm going to pull in the counts matrix to make sure that GFP was present in there to begin with (which I believe it should be).

The counts matrix from geo is in misc/SRR26705626.h5

```{r check-geo-matrix}
og_sobj <- tenx_load_qc(h5_file = "misc/SRR26705626.h5",
                        violin_plot = FALSE)

feature_hist(og_sobj, features = c("nCount_RNA", "percent.mt"))

#using same cutoffs as before
og_sobj <- subset(og_sobj,
               nCount_RNA >= 1200 &
               nCount_RNA <= 50000 &
               percent.mt < 10) %>%
    process_seurat()

r_dim_plot(og_sobj)

r_feature_plot(og_sobj, features = c("Gfp", "GFP", "gfp"))
grep("GFP", rownames(og_sobj), ignore.case = TRUE, value = TRUE)

#annotate tumor cells using their markers


#Seems like I'm using the wrong name for GFP, so I'm going to run find markers and see what has the biggest difference in pct expressing for tumor
tumor_feats <- c("Crabp1", "Pebp1", "Tm4sf1", "Igfbp4", "Sdc2", "Emp2", "Ank")
r_dim_plot(og_sobj) | r_feature_plot(og_sobj, tumor_feats, pt.size = 0.5)

#8 4 and 10 look like our tumor cells
og_sobj$tumor <- ifelse(og_sobj$seurat_clusters %in% c(4, 8, 10),
                        "tumor",
                        "not_tumor")

Idents(og_sobj) <- og_sobj$tumor
og_genes <- FindAllMarkers(og_sobj, only.pos = TRUE)
head(og_genes)
#Sort for positive in tumors and highly expressed in tumors
og_genes %>%
    subset(cluster == "tumor" &
           p_val_adj < 0.05) %>%
    select(avg_log2FC,
           pct.1,
           pct.2,
           gene) %>%
    mutate(pct.diff = pct.1 - pct.2) %>%
    arrange(desc(pct.diff)) %>%
    head(n = 50)

```