## Validate with B6 and Balbc
- S0066 - B6 lung
- S0075 - Balbc lung
```{r mouse_strain_make_sobj}
#| dependson: [read_raw_mouse, scanbit_run_mouse]
merged_norm_mouse <-
    qs::qread("output/seurat_objects/merged/merged_mouse.qs") %>%
    JoinLayers() %>%
    subset(Sample_ID %in% c("S0066", "S0074")) %>%
    process_seurat()

qs::qsave(
    merged_norm_mouse,
    "output/seurat_objects/merged/merged_norm_mouse.qs"
)

merged_norm_mouse$cell_group %>%
    unique() %>%
    write_lines("output/snv_processing/norm_mouse/keep_clusters.txt")

norm_mouse_dimplot <-
    DimPlot(
        merged_norm_mouse,
        group.by = "cell_group",
        label = TRUE,
        label.box = TRUE,
        repel = TRUE
    ) +
    NoLegend()

ggsave(
    "output/figures/norm_mouse_dimplot_clusters.pdf",
    plot = norm_mouse_dimplot,
    width = 10,
    height = 10
)
```

## Filter the merged bcf file for just these samples
```{bash mouse_strain_filter_bcfs}
#| dependson: [scanbit_run_mouse]
conda activate scanBit_xkcd_1337

for cutoff in 5 10 20 30
do
bcftools \
    view \
    -S output/snv_processing/norm_mouse/keep_clusters.txt \
    output/snv/mouse/mergedmouse_snvs_c${cutoff}.bcf \
    -o output/snv_processing/norm_mouse/mergedmouse_snvs_c${cutoff}_filtered.bcf
done
```

## Make trees directly from merged bcfs
This is so the figures and downstream analyses only have B6 and BALB/c
```{r mouse_strain_trees}
#| dependson: [mouse_strain_make_sobj]
min_snvs_per_cluster <- 250
n_bootstraps <- 10000
bootstrap_cutoff <- 0.95

dist_worked <-
    parallel::mclapply(
        c("5", "10", "20", "30"),
        mc.cores = 4,
        mc.preschedule = FALSE,
        function(cutoff) {
            group_clusters_by_dist(
                merged_bcf_file = paste0(
                    "output/snv_processing/norm_mouse/mergedmouse_snvs_c",
                    cutoff,
                    "_filtered.bcf"
                ),
                output_file = paste0(
                    "output/snv_processing/norm_mouse/mergedmouse_snvs_c",
                    cutoff,
                    "_groups.txt"
                ),
                min_snvs_per_cluster = min_snvs_per_cluster,
                n_bootstraps = n_bootstraps,
                bootstrap_cutoff = bootstrap_cutoff,
                tree_figure_file = paste0(
                    "output/snv_processing/norm_mouse/mergedmouse_snvs_c",
                    cutoff,
                    "_tree.pdf"
                ),
                verbose = FALSE,
                other_job_header_options = c(
                    "--time=8:00:00",
                    "--partition=himem,general"
                ),
                other_batch_options = c(
                    "ml purge",
                    "ml miniforge3",
                    'eval "$(conda shell.bash hook)"'
                ),
                sge_q = "",
                sge_parallel_environment = ""
            )
        }
    )

```


##
I want to look over the merged bcfs and quantify the proportion of positions covered by n reads has a variant call.

Then I want to compare these values between the flex data, where any variant call is artifactual and the B6 data, where an unknown proportion of variant calls at any depth are artifactual. This will allow me to estimate the false positive rate for SNV calls in the pipeline.

```{r}
merged_sobj_mouse_ours <-
    qs::qread("output/seurat_objects/merged/merged_mouse.qs")

# Grab just data from our lab - excludes flex data
merged_sobj_mouse_ours <-
    merged_sobj_mouse_ours[, grepl("^S0", merged_sobj_mouse_ours$Sample_ID)]

merged_sobj_mouse_ours$cell_group <-
    paste0(
        merged_sobj_mouse_ours$cluster_label,
        "_c",
        merged_sobj_mouse_ours$individ_clusters
    )
```

```{r mouse_scanbit}
cell_barcode_group_table <-
    merged_sobj_mouse_ours@meta.data %>%
    select(cell_group, cell_barcode, bam_file) %>%
    as_tibble()

snp_tree <-
    scanBit::get_snp_tree(
        cellid_bam_table = cell_barcode_group_table,
        ploidy = "mm10",
        output_dir = "output/snv/mouse/",
        temp_dir = "output/snv/mouse/tempdir",
        output_base_name = "mouse_ours",
        ref_fasta = "/reference/mus_musculus/mm10/ucsc_assmebly/illumina_download/Sequence/WholeGenomeFasta/genome.fa",
        min_depth = c(1, 5, 10, 20, 30),
        sbatch_base = "our_mouse_sbatch",
        min_snvs_per_cluster = 500,
        n_bootstraps = 10000,
        max_prop_missing_at_site = 0.9,
        cleanup = TRUE,
        other_sbatch_options = "--time=24:00:00"
    )
```

## Get merged bcf with all sites
To estimate the proportion of variant calls, I set temp_dir to a folder in output and specified cleanup = FALSE. This makes the shell scripts to generate the data write to an accessible folder so I can tweak the shell script manually. I got rid of the `-i 'GT[*]="alt"' ` filter and wrote to a new bcf file.

I then manually deleted the large intermediate files

```{bash mouse_split_bcf}
conda activate scanBit_xkcd_1337
# S0066, S0259, S0282 are B6 or B6 derived F420
bcftools view -h output/snv/mouse/mergedmouse_ours_c5.bcf \
    | grep "^#CHROM" \
    | perl -pe 's/\t/\n/g' \
    | grep "S0066\|S0067\|S0068\|S0200\|S0259\|S0269\|S0282\|S0291" \
    > output/snv/mouse/b6_group_clusters.txt

bcftools view \
    -O b \
    -o output/snv/mouse/mergedmouse_ours_c5_B6_group.bcf \
    -S output/snv/mouse/b6_group_clusters.txt \
    output/snv/mouse/mergedmouse_ours_c5.bcf

bcftools index output/snv/mouse/mergedmouse_ours_c5_B6_group.bcf

# S0074, S0105, S0122 are Balbc or Balbc derived K7M2
for cutoff in 1 5 10 20 30
do
    bcftools view -h output/snv/mouse/mergedmouse_ours_c${cutoff}.bcf \
        | grep "^#CHROM" \
        | perl -pe 's/\t/\n/g' \
        | grep "S0074\|S0075\|S0105\|S0122\|S0190\|S0201\|S0238" \
        > output/snv/mouse/balb_group_clusters${cutoff}.txt

    bcftools view \
        -O b \
        -o output/snv/mouse/mergedmouse_ours_c${cutoff}_balbc_group.bcf \
        -S output/snv/mouse/balb_group_clusters${cutoff}.txt \
        output/snv/mouse/mergedmouse_ours_c${cutoff}.bcf

    bcftools index output/snv/mouse/mergedmouse_ours_c${cutoff}_balbc_group.bcf

done
```

```{r mouse_subtree}
just_b6 <-
    merged_sobj_mouse_ours %>%
    subset(Sample_ID %in% c(
        "S0066", "S0067", "S0068", "S0200", "S0259", "S0269", "S0282", "S0291"
    )) %>%
    process_seurat()

other_group <- c("S0200_c1", "S0291_c2", "S0259_c0", "S0200_c0", "S0067_c6", "S0068_c5", "S0269_c10", "S0269_c1")

just_b6$f420_group <-
    just_b6$cell_group %in% other_group

png(
    "output/figures/b6_groups_dimplot.png",
    width = 4000,
    height = 8000,
    res = 300
)
(DimPlot(
    just_b6,
    group.by = c("f420_group", "cell_group"),
    label = TRUE,
    label.box = TRUE,
    repel = TRUE,
    ncol = 1
) +
    NoLegend()) +
    (DimPlot(
        just_b6,
        group.by = "cell_type",
        label = TRUE,
        label.box = TRUE,
        repel = TRUE,
        ncol = 1
    ) +
        NoLegend())

dev.off()

scanBit::group_clusters_by_dist(
    merged_bcf_file = "output/snv/mouse/mergedmouse_ours_c5_B6_group.bcf",
    output_file = "output/snv/mouse/mergedmouse_ours_c5_B6_group.tsv",
    min_snvs_per_cluster = 500,
    max_prop_missing_at_site = 0.9,
    tree_figure_file = "output/snv/mouse/mergedmouse_ours_c5_B6_group_tree.png",
    temp_dir = ".",
    n_bootstraps = 10000,
    other_sbatch_options = "--time=24:00:00"
)

scanBit::group_clusters_by_dist(
    merged_bcf_file = "output/snv/mouse/mergedmouse_ours_c10_balbc_group.bcf",
    output_file = "output/snv/mouse/mergedmouse_ours_c10_balbc_group.tsv",
    min_snvs_per_cluster = 500,
    max_prop_missing_at_site = 0.9,
    tree_figure_file = "output/snv/mouse/mergedmouse_ours_c10_balbc_group_tree.png",
    temp_dir = ".",
    n_bootstraps = 10000,
    other_sbatch_options = "--time=24:00:00"
)
```
