Before I download all the available data, I'm going to test it on the control samples from GSE247228.

The sra Run numbers are as follows:
    - SRR26705626
    - SRR26705627
    - SRR26705629

## Get FASTQ Files

I'll be using SRA-Toolkit/2.10.8 for this. Specifically, the `fasterq-dump` function.

### GSE247228

```{bash get-computing-power}
srun_l -c 12 -p himem --x11 --pty bash
```

```{bash get-fastqs}
ml SRA-Toolkit/2.10.8
fasterq-dump --outdir input/fastqs/GSE247228 --progress SRR26705626 SRR26705627 SRR26705629 SRR26705628 SRR26705630 SRR26705631
```

### GSE179077

Fluorescent UnaG in hypoxic glioblastoma cells

```{bash get-unag-fastqs}
fasterq-dump --outdir input/fastqs/GSE179077 --progress SRR29428580 SRR29428581 SRR29428582 SRR29428583 SRR29428584 SRR29428585 SRR29428586 SRR29428587
```

## Update Reference Genome

I need to add the fluorescent reporter sequence to our reference genome before aligning our fastqs. I'll be adding it to our 10x-mm10 reference genome.

### eGFP

I will first need to create a file named GFP.fa containing the nucleotide sequence for our GFP protein.

```{bash make-new-ref}
#get number of bases in GFP.fa
grep -v "^>" custom_ref/eGFP/eGFP.fa | tr -d "\n" | wc
#720 bases

#lets make a custom gtf for gfp; \t helps us make new columns that are needed in the gtf; we need both the exon and transcript lines 
echo -e 'eGFP\tunknown\texon\t1\t720\t.\t+\t.\tgene_id "eGFP"; transcript_id "eGFP"; gene_biotype "proetin_coding";\n'\
'eGFP\tunknown\ttranscript\t1\t720\t.\t+\t.\tgene_id "eGFP"; transcript_id "eGFP"; gene_biotype "proetin_coding";\n'\
'eGFP\tunknown\tgene\t1\t720\t.\t+\t.\tgene_id "eGFP"; gene_name "eGFP"; gene_biotype "proetin_coding";' > custom_ref/eGFP/eGFP.gtf

#copy mm10 reference FASTA over (so we can modify it and not screw up the og)
cp /home/gdrobertslab/lab/GenRef/10x-mm10/fasta/genome.fa custom_ref/eGFP/genome.fa
chmod 744 custom_ref/eGFP/*

#append GFP.fa to end of mm10 genome
cat custom_ref/eGFP/eGFP.fa >> custom_ref/eGFP/genome.fa

#check that GFP "chromosome" is present
grep ">" custom_ref/eGFP/genome.fa

#copy over original gtf and gunzip
cp /home/gdrobertslab/lab/GenRef/10x-mm10/genes/genes.gtf.gz custom_ref/eGFP/.
chmod 744 custom_ref/eGFP/*
gunzip custom_ref/eGFP/genes.gtf.gz

#append GFP gtf to mm10 gene file
cat custom_ref/eGFP/eGFP.gtf >> custom_ref/eGFP/genes.gtf

#Time to make our new reference using cellranger mkref
cellranger mkref --output-dir custom_ref/eGFP/demo_eGFP \
    --genome demo_eGFP \
    --fasta custom_ref/eGFP/genome.fa \
    --genes custom_ref/eGFP/genes.gtf
```

Now I'm going to make a genome containing the hypoxic reporter element plus unag.

```{bash make-unag-ref}
#making a folder within custom ref for this, otherwise it'll get too cluttered
mkdir custom_ref/unag/

grep -v "^>" custom_ref/unag/hre_unag.fa | tr -d "\n" | wc
#8646 bases


echo -e 'HRE-dUnaG\tunknown\texon\t1\t8646\t.\t+\t.\tgene_id "HRE-dUnaG"; transcript_id "HRE-dUnaG"; gene_biotype "proetin_coding";\n'\
'HRE-dUnaG\tunknown\ttranscript\t1\t8646\t.\t+\t.\tgene_id "HRE-dUnaG"; transcript_id "HRE-dUnaG"; gene_biotype "proetin_coding";\n'\
'HRE-dUnaG\tunknown\tgene\t1\t8646\t.\t+\t.\tgene_id "HRE-dUnaG"; gene_name "HRE-dUnaG"; gene_biotype "proetin_coding";' > custom_ref/unag/hre_unag.gtf

#Going to append the gtf and fasta to the mm10
cp /home/gdrobertslab/lab/GenRef/10x-mm10/genes/genes.gtf.gz custom_ref/unag/genes.gtf.gz

cp /home/gdrobertslab/lab/GenRef/10x-mm10/fasta/genome.fa custom_ref/unag/genome.fa

chmod 744 custom_ref/unag

gunzip custom_ref/unag/genes.gtf.gz

#cat gtf and fasta for hre-unag to mm10 reference files
cat custom_ref/unag/hre_unag.fa >> custom_ref/unag/genome.fa
cat custom_ref/unag/hre_unag.gtf >> custom_ref/unag/genes.gtf

#make reference
ml cellranger
cellranger mkref --output-dir custom_ref/unag/unag_ref \
    --genome unag_ref \
    --fasta custom_ref/unag/genome.fa \
    --genes custom_ref/unag/genes.gtf \
    --jobmode=local \
    --localcores=8
```

## Run cellranger count

### On eGFP data From GSE247228

```{r mk-bams}
# We need to convert our fastq names to a cellranger compatible format

for sra in SRR26705626 SRR26705627 SRR26705628 SRR26705629 SRR26705630 SRR26705631
do
    mkdir input/fastqs/${sra}
    mv input/fastqs/GSE247228/${sra}_1.fastq \
        input/fastqs/GSE247228/${sra}/${sra}_S1_R1_001.fastq
    mv input/fastqs/GSE247228/${sra}_2.fastq \
        input/fastqs/GSE247228/${sra}/${sra}_S1_R2_001.fastq
done

#run cellranger count!
sbatch cellranger_count_egfp.sh
```

### On unag data from 

## Look at New Object

```{r}
sobj <- tenx_load_qc("counts/SRR26705629/outs/filtered_feature_bc_matrix", violin_plot = FALSE) 
sobj <- process_seurat(sobj)
r_feature_plot(sobj, features = c("eGFP"))
sum(rownames(sobj) == "eGFP")

# feature_hist(sobj,
#              features = c("nCount_RNA", "percent.mt"),
#              cutoff_table =
#     tribble(~feature, ~min_val, ~max_val,
#             "nCount_RNA", 1200, 80000,
#             "percent.mt", 0, 10))

# sobj <- subset(sobj,
#                nCount_RNA >= 1200 &
#                nCount_RNA <= 50000 &
#                percent.mt < 10) %>%
#     process_seurat()

```

Weirdly, I'm not seeing any GFP expression in any of the cells. I'm going to pull in the counts matrix to make sure that GFP was present in there to begin with (which I believe it should be).

The counts matrix from geo is in misc/SRR26705626.h5

```{r check-geo-matrix}
og_sobj <- tenx_load_qc(h5_file = "misc/SRR26705626.h5",
                        violin_plot = FALSE)

feature_hist(og_sobj, features = c("nCount_RNA", "percent.mt"))

#using same cutoffs as before
og_sobj <- subset(og_sobj,
               nCount_RNA >= 1200 &
               nCount_RNA <= 50000 &
               percent.mt < 10) %>%
    process_seurat()

r_dim_plot(og_sobj)

r_feature_plot(og_sobj, features = c("Gfp", "GFP", "gfp"))
grep("GFP", rownames(og_sobj), ignore.case = TRUE, value = TRUE)

#annotate tumor cells using their markers


#Seems like I'm using the wrong name for GFP, so I'm going to run find markers and see what has the biggest difference in pct expressing for tumor
tumor_feats <- c("Crabp1", "Pebp1", "Tm4sf1", "Igfbp4", "Sdc2", "Emp2", "Ank")
r_dim_plot(og_sobj) | r_feature_plot(og_sobj, tumor_feats, pt.size = 0.5)

#8 4 and 10 look like our tumor cells
og_sobj$tumor <- ifelse(og_sobj$seurat_clusters %in% c(4, 8, 10),
                        "tumor",
                        "not_tumor")

Idents(og_sobj) <- og_sobj$tumor
og_genes <- FindAllMarkers(og_sobj, only.pos = TRUE)
head(og_genes)
#Sort for positive in tumors and highly expressed in tumors
og_genes %>%
    subset(cluster == "tumor" &
           p_val_adj < 0.05) %>%
    select(avg_log2FC,
           pct.1,
           pct.2,
           gene) %>%
    mutate(pct.diff = pct.1 - pct.2) %>%
    arrange(desc(pct.diff)) %>%
    head(n = 50)

```